---
title: "Pathogenic Splice site identification: developing filters to be used"
output: html_notebook
---

Import libraries
```{r}
suppressPackageStartupMessages(library(Homo.sapiens))
```


Functions
```{r}
#Function to convert bed--> genomic ranges --> add gene IDs --> select gene ID of interest
GetJunctionGeneID<- function(bed, genes, GeneIDs) {
  #Set column names
  colnames(bed)<- c("Chrom", "Start", "End", "Name", "Score", "Strand", "ThickStart", "ThickEnd","ItemRGB", "BlockCount", "BlockSizes", "BlockStarts")
  #Convert bed file to genomic ranges, add score as metadata column
  GR<- makeGRangesFromDataFrame(bed)
  values(GR)<- DataFrame(Score=bed$Score, Name=bed$Name)
  
  #Get overlaps of GRanges bed info and all known genes
  ranges<- subsetByOverlaps(GR,genes)
  
  #Get the geneID names, associate back with the overlap GRanges object
  hits<- findOverlaps(GR,genes)
  geneid <- CharacterList(split(genes$gene_id[subjectHits(hits)],queryHits(hits)))
  mcols(ranges) <- DataFrame(mcols(ranges), geneid)
  
  Junctions_GeneIDs<- as.data.frame(ranges)
  GeneJunctions_GeneIDs<- Junctions_GeneIDs[Junctions_GeneIDs$geneid %in% GeneIDs,]
  return(GeneJunctions_GeneIDs)
}

#Write function that filters junctions based on pre-built control list of junctions + depth
FilterJunctions<- function(ControlJunctions, SampleJunctions, DepthFilter){
  ControlJunctionsInSample<- SampleJunctions$Name %in% ControlJunctions$Name
  UniqueTTNJunctions<- SampleJunctions[!ControlJunctionsInSample,]
  UniqueSample_Depth<- UniqueTTNJunctions[UniqueTTNJunctions$Score>DepthFilter,]
  print(nrow(UniqueSample_Depth))
  return(UniqueSample_Depth)
}
```


#Building filtering mechanism

Import bed files: all 5 controls + 2 samples with TTN splice variants we already found
```{r}
#For novel tx alignment
wd<- "//Cifs2/rcdata$/UKRNASeq/RawFiles/Bed/"
#For Original alignment
wd<- "//Cifs2/rcdata$/UKRNASeq/PathogenicSpliceSiteAnalysis/Junctions.bed/"

bed_s2<- read.table(paste(wd, "2.junctions.bed", sep=""),sep ="\t", header=FALSE, stringsAsFactors=FALSE)
bed_s9<- read.table(paste(wd, "9.junctions.bed", sep=""),sep ="\t", header=FALSE, stringsAsFactors=FALSE)

bed_control1<- read.table(paste(wd, "UK-Control-1.junctions.bed", sep=""),sep ="\t", header=FALSE, stringsAsFactors=FALSE)
bed_control2<- read.table(paste(wd, "UK-Control-2.junctions.bed", sep=""),sep ="\t", header=FALSE, stringsAsFactors=FALSE)
bed_control3<- read.table(paste(wd, "UK-Control-3.junctions.bed", sep=""),sep ="\t", header=FALSE, stringsAsFactors=FALSE)
bed_control4<- read.table(paste(wd, "UK-Control-4.junctions.bed", sep=""),sep ="\t", header=FALSE, stringsAsFactors=FALSE)
bed_control5<- read.table(paste(wd, "UK-Control-5.junctions.bed", sep=""),sep ="\t", header=FALSE, stringsAsFactors=FALSE)
```


Get TTN junctions for all 7 samples
```{r}
#get all known genes as Genomic ranges object
genes<-genes(TxDb.Hsapiens.UCSC.hg19.knownGene)

TTN_Junctions_Sample2<- GetJunctionGeneID(bed_s2, genes, "7273") #775 junctions
TTN_Junctions_Sample9<- GetJunctionGeneID(bed_s9, genes, "7273") #591 junctions

TTN_Junctions_Control1<- GetJunctionGeneID(bed_control1, genes, "7273") #668 junctions
TTN_Junctions_Control2<- GetJunctionGeneID(bed_control2, genes, "7273") #727 junctions
TTN_Junctions_Control3<- GetJunctionGeneID(bed_control3, genes, "7273") #633 junctions
TTN_Junctions_Control4<- GetJunctionGeneID(bed_control4, genes, "7273") #624 Junctions
TTN_Junctions_Control5<- GetJunctionGeneID(bed_control5, genes, "7273") #625 junctions

```

Get list of all TTN junctions found in sample 2 or sample 9 but NOT found in any of the controls
```{r}
#List of all TTN junctions in controls
ControlTTNJunctions<- unique(c(TTN_Junctions_Control1$Name, TTN_Junctions_Control2$Name, TTN_Junctions_Control3, TTN_Junctions_Control4, TTN_Junctions_Control5))

#junctions in sample 2 that are not in controls
ControlJunctionsInSample2<- TTN_Junctions_Sample2$Name %in% ControlTTNJunctions
Sample2UniqueTTNJunctions<- TTN_Junctions_Sample2[!ControlJunctionsInSample2,] #127 unique sample 2 junctions

#junctions in sample 9 that are not in controls
ControlJunctionsInSample9<- TTN_Junctions_Sample9$Name %in% ControlTTNJunctions
Sample9UniqueTTNJunctions<- TTN_Junctions_Sample9[!ControlJunctionsInSample9,] #46 unique sample 9 junctions
```

Additional filtering of the two lists of unique junctions
```{r}
#At least 2 reads to support junction
UniqueSample2_Depth<- Sample2UniqueTTNJunctions[Sample2UniqueTTNJunctions$Score>2,] #53 pass threshold
UniqueSample9_Depth<- Sample9UniqueTTNJunctions[Sample9UniqueTTNJunctions$Score>2,] #19 pass threshold

#Not found in the other sample
SuperUniqueSample2<- UniqueSample2_Depth[!(UniqueSample2_Depth$Name %in% TTN_Junctions_Sample9),] #Still 53
SuperUniqueSample9<- UniqueSample9_Depth[!(UniqueSample9_Depth$Name %in% TTN_Junctions_Sample2),] #still 19
```

```{r}
#For sample 9: our "positive control" splice site chr2:179496028-179497001 is not showing up in this list.

#It's removed when we filter by controls. So what controls have this splice site?
TTN_Junctions_Control1[grep("chr2:179496028-179497001", TTN_Junctions_Control1$Name),] #4
TTN_Junctions_Control2[grep("chr2:179496028-179497001", TTN_Junctions_Control2$Name),] #0
TTN_Junctions_Control3[grep("chr2:179496028-179497001", TTN_Junctions_Control3$Name),] #0
TTN_Junctions_Control4[grep("chr2:179496028-179497001", TTN_Junctions_Control4$Name),] #0
TTN_Junctions_Control5[grep("chr2:179496028-179497001", TTN_Junctions_Control5$Name),] #0
#Only control 1 has it, and at very low levels. When filtering junctions because they're found in controls, should either set a threshold to only count when controls have higher depth (i.e., at least 10 reads supporting) or when found in more than 1 control. Either of these thresholds would have resulted in this junction being discovered.

#In samples 2 and 9
TTN_Junctions_Sample2[grep("chr2:179496028-179497001", TTN_Junctions_Sample2$Name),] #8
TTN_Junctions_Sample9[grep("chr2:179496028-179497001", TTN_Junctions_Sample9$Name),] #605

#For sample 2: our "positive control" splice site is at the region chr2:179,602,998-179,603,177 chr2:179602817-179604260
#In this region, there should be several splice variants.
#We did actually find this:
sample2_Pos<- makeGRangesFromDataFrame(UniqueSample2_Depth)
values(sample2_Pos)<- DataFrame(Score=UniqueSample2_Depth$Score, Name=UniqueSample2_Depth$Name)

PosControl<-GRanges(seqnames =Rle(c("chr2"), c(2)),
                    ranges =IRanges(179602817, end = 179604260),
                    strand =Rle(strand(c("-")),c(1)))

Sample2_AtPos<-as.data.frame(subsetByOverlaps(sample2_Pos,PosControl)) #11 junctions that OVERLAP the region of interest. 

Sample2_AtPos
#Looking at these junctions in IGV, the last 3 are actually at this junction, not spreading way over across this junction
#chr2:179603079-179603867
#chr2:179603167-179603867
#chr2:179603229-179603867

#So: these 3 junctions are ones we wanted to end up with. Looking at the bed file in IGV, I also can't see any that we would want to mark as interesting in this area but didn't make it through the filter
```

So for sample 9, we need to decrease the stringency of our control filters to allow for at least 1 control sample having a couple reads supporting that junction, and still mark it interesting



For sample 2, the question is: how do we increase the filter to get rid of all those other ones we're not interested in?

Some ideas:
1) Don't look at junctions that span too far. Kind of arbitrary, and may not be the case in all. But if a junction is spanning >10kb, it's probably just random mapping
2) Increase the score filters- junctions with fewer than 10 reads aren't likely to be interesting, unless we're REALLY affecting expression by messing with splicing. So may not want to do this.
3) Set a similar filter as in the paper- one side of the junction has to be canonical. We could define canonical as occurring in the control set a bunch of times- like at least 100 depth
4) Compare it to the other samples- like paper, that junction has to have highest depth in that sample, by at least a factor of 2


```{r}
#Discovering positive control splice site in sample 9: decreasing stringency of control filters

#Take list of all junctions found in controls --> to data.frame with 5 columns (1 for each control sample) for each junction. Include the number of reads that support that junction for each sample, entering a 0 if it doesn't appear
Merge1_2<- merge(TTN_Junctions_Control1[,6:7],TTN_Junctions_Control2[,6:7], all=TRUE, by= "Name")
Merge3<- merge(TTN_Junctions_Control3[,6:7], Merge1_2, all=TRUE,by="Name")
Merge4<- merge(TTN_Junctions_Control4[,6:7], Merge3, all=TRUE,by="Name")
Merge5<- merge(TTN_Junctions_Control5[,6:7], Merge4, all=TRUE,by="Name")
colnames(Merge5)<- c("Name", "Score_Control5", "Score_Control4", "Score_Control3", "Score_Control2", "Score_Control1")

#Convert NA to 0
Merge5[is.na(Merge5)]<- 0

#Add column with maximum score for each junction
Merge5$Max<- rowMax(as.matrix(Merge5[,2:6]))

#Add column with Number of samples that have that junction
Merge5$SamplesWithJunction<- rowSums(Merge5[,2:6]>0)

#Get list of junctions to NOT look at in sample 9: Appears in >1 control samples + supported by > 10 reads
ControlJunctions_Filtered<- Merge5[Merge5$Max>10,]
ControlJunctions_Filtered<- ControlJunctions_Filtered[ControlJunctions_Filtered$SamplesWithJunction>1,]
#Down to 464 junctions- removed 434 junctions from "controls" list

#junctions in sample 9 that are not in filtered control list
ControlJunctionsInSample9<- TTN_Junctions_Sample9$Name %in% ControlJunctions_Filtered$Name
Sample9UniqueTTNJunctions<- TTN_Junctions_Sample9[!ControlJunctionsInSample9,] #160 sample 9 junctions NOT in filtered controls list

#Filtering unique sample 9 junctions by depth
UniqueSample9_Depth<- Sample9UniqueTTNJunctions[Sample9UniqueTTNJunctions$Score>2,] #56 pass threshold
UniqueSample9_Depth<- Sample9UniqueTTNJunctions[Sample9UniqueTTNJunctions$Score>20,] #1 passes threshold

UniqueSample9_Depth[grep("chr2:179496028-179497001", UniqueSample9_Depth$Name),]

#With these new control filters, this "positive control" junction for sample 9 can now be seen in our sample 9 "shortlist"
#Also: this junction is supported by 605 reads in sample 9. The next highest score in our shortlist is 11. If we do more stringent depth filtering to narrow down shortlist even more, (score>20), we're left with only 1 junction- our positive control.


#Filtering sample 2 based on filterd control list --> more stringent depth

#junctions in sample 2 that are not in filtered control list
ControlJunctionsInSample2<- TTN_Junctions_Sample2$Name %in% ControlJunctions_Filtered$Name
Sample2UniqueTTNJunctions<- TTN_Junctions_Sample2[!ControlJunctionsInSample2,] #319 sample 2 junctions NOT in filtered controls list

#Filtering unique sample 2 junctions by depth
UniqueSample2_Depth<- Sample2UniqueTTNJunctions[Sample2UniqueTTNJunctions$Score>2,] #139 pass threshold
UniqueSample2_Depth<- Sample2UniqueTTNJunctions[Sample2UniqueTTNJunctions$Score>20,] #5 pass threshold

#Only 5 pass threshold. So just filtering based on occurence in controls and depth might be a quick and dirty way to find interesting titin junctions to look at more closely in IGV.

#Repeating this analysis for the other samples.
```