---
title: "DESeq2: Final"
output: html_notebook
---

#Running DESeq2 analysis for UK RNASeq
Final, simplified analysis

libraries:
```{r}
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(pheatmap))
```

Importing data:
```{r}
#set path to file
wd<- "//Cifs2/rcdata$/UKRNASeq/"
countPath<- paste(wd, "RawCounts/CountData.csv", sep = "")

#set sample groups
sampleGroup<- c("pos", "pos", "neg", "pos", "neg", "neg", "neg", "pos", "neg", "neg", "neg", "neg", "neg", "pos", "neg", "neg", rep("control", 5))

#read in count data, format, get rid of NA data, get rid of any genes where all counts are 0
countsInput<- read.csv(countPath)
rownames(countsInput)<- countsInput$X
countsInput<- countsInput[,-1]
colnames(countsInput)<- gsub(".counts.*", "", colnames(countsInput))
counts<- na.omit(countsInput)
counts<- counts[rowSums(counts)>0,]

#create metadata
sample<- colnames(countsInput)
meta<- data.frame(sample, sampleGroup)
rownames(meta)<- meta$sample
```

Creating DESeq2 object and running DESeq2
```{r}
#check that counts/meta data has samples in same order
all(rownames(meta) %in% colnames(counts))
all(rownames(meta) == colnames(counts))

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=counts, 
                             colData=meta, 
                             design=~sampleGroup)

#define levels- so that controls will always be used as reference group
dds$sampleGroup<- relevel(dds$sampleGroup, ref= "control")

#run DE
dds<- DESeq(dds)

#normalize dds for later use. Only need to run this once.
rld<- rlog(dds)
write.csv(assay(rld), paste(wd, "rldNorm_AfterDESeq2.csv", sep=''))
```


#Assess data quality
```{r}
#PCA
vsd <- vst(dds, blind=FALSE)
PCAData<-plotPCA(vsd, intgroup="sampleGroup", returnData=TRUE)
#sample 3: PC1 = 3.95 PC2= -1.07
plotPCA(vsd, intgroup="sampleGroup")
```

Get results for all 3 groups
```{r}
#function that takes dds object and desired contrasts as input, outputs ordered results
ContrastDGE<- function(contrasts, dds){
  ContrastsUse<- c("sampleGroup", contrasts)
  ContrastsResults<- na.omit(results(dds, contrast=ContrastsUse))
  resOrdered<- ContrastsResults[order(ContrastsResults$padj),]
  return(resOrdered)
}

#Extract comparisons for 3 desired groups
DGERes_controlpos<- ContrastDGE(c("control", "pos"), dds)
DGERes_controlneg<- ContrastDGE(c("control", "neg"), dds)
DGERes_posneg<- ContrastDGE(c("pos", "neg"), dds)

nrow(DGERes_controlpos[DGERes_controlpos$padj<0.05,])
nrow(DGERes_controlneg[DGERes_controlneg$padj<0.05,])
nrow(DGERes_posneg[DGERes_posneg$padj<0.05,])
```

Write to csv
```{r}
write.csv(DGERes_controlpos, paste(wd, "DESeq2/Final/", "DGERes_controlvspos.csv", sep = ""))
write.csv(DGERes_controlneg, paste(wd, "DESeq2/Final/", "DGERes_controlvsneg.csv", sep = ""))
write.csv(DGERes_posneg, paste(wd, "DESeq2/Final/", "DGERes_posvsneg.csv", sep = ""))
```


#Clustering DESeq2 results in heatmap

Select only significant DEGs
```{r}
posSig<- DGERes_controlpos[DGERes_controlpos$padj<= 0.05,]
negSig<- DGERes_controlneg[DGERes_controlneg$padj<= 0.05,]
posnegSig<- DGERes_posneg[DGERes_posneg$padj<= 0.05,]
```

Select only DEGs with logfoldchange above 1
```{r}
Above1Pos<- posSig[abs(posSig$log2FoldChange)>=1,]
Above1Neg<- negSig[abs(negSig$log2FoldChange)>=1,]
Above1PosNeg<- posnegSig[abs(posnegSig$log2FoldChange)>=1,]

subtractFromMin<-0
subtractFromMax<- 2.5
lengthOut<- 75
```


Plotting heatmaps: multiple options

#Option #1: Using all sig genes
```{r}
topgenes<- unique(c(rownames(negSig), rownames(posSig)))
topgenes<- unique(c(topgenes, rownames(posnegSig)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,9
```


#Option #2: Using top 50 sig genes from each
```{r}
topgenes<- unique(c(rownames(negSig[1:50,]), rownames(posSig[1:50,])))
topgenes<- unique(c(topgenes, rownames(posnegSig[1:50,])))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,3,5,9,17
```



#Option #3: Using top 100 sig genes from pos/neg vs control and all of pos vs neg
```{r}
topgenes<- unique(c(rownames(negSig[1:100,]), rownames(posSig[1:100,])))
topgenes<- unique(c(topgenes, rownames(posnegSig)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,3,5,9,17
```
#Option #4: Using top 75 sig genes from pos/neg vs control and all of pos vs neg
```{r}
topgenes<- unique(c(rownames(negSig[1:75,]), rownames(posSig[1:75,])))
topgenes<- unique(c(topgenes, rownames(posnegSig)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,3,5,9,17
```

#Option #4: Using top 200 sig genes from pos/neg vs control and all of pos vs neg
```{r}
topgenes<- unique(c(rownames(negSig[1:200,]), rownames(posSig[1:200,])))
topgenes<- unique(c(topgenes, rownames(posnegSig)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,9
```

#Option #4: Using top 150 sig genes from pos/neg vs control and all of pos vs neg
```{r}
topgenes<- unique(c(rownames(negSig[1:150,]), rownames(posSig[1:150,])))
topgenes<- unique(c(topgenes, rownames(posnegSig)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,9
```

#Option #4: Using top 125 sig genes from pos/neg vs control and all of pos vs neg
```{r}
topgenes<- unique(c(rownames(negSig[1:125,]), rownames(posSig[1:125,])))
topgenes<- unique(c(topgenes, rownames(posnegSig)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,5,9,3,17
```

#Option #4: Use sig genes with fold change above 1
```{r}
topgenes<- unique(c(rownames(Above1Neg), rownames(Above1Pos)))
topgenes<- unique(c(topgenes, rownames(Above1PosNeg)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: controls/1,2,9
```

#Option #5: Use sig genes with fold change above .5
```{r}
Above1Pos<- posSig[abs(posSig$log2FoldChange)>=.5,]
Above1Neg<- negSig[abs(negSig$log2FoldChange)>=.5,]
Above1PosNeg<- posnegSig[abs(posnegSig$log2FoldChange)>=.5,]

topgenes<- unique(c(rownames(Above1Neg), rownames(Above1Pos)))
topgenes<- unique(c(topgenes, rownames(Above1PosNeg)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: controls/1,2,9
```


#Option #5: Use sig genes with fold change above .1
```{r}
Above1Pos<- posSig[abs(posSig$log2FoldChange)>=.1,]
Above1Neg<- negSig[abs(negSig$log2FoldChange)>=.1,]
Above1PosNeg<- posnegSig[abs(posnegSig$log2FoldChange)>=.1,]

topgenes<- unique(c(rownames(Above1Neg), rownames(Above1Pos)))
topgenes<- unique(c(topgenes, rownames(Above1PosNeg)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: controls/1,2,9
```


Conclusion: using a log2 fold change cutoff isn't super useful; doesn't seem to improve clustering.




