---
title: "DESeq2: Final"
output: html_notebook
---

#Running DESeq2 analysis for UK RNASeq
Final, simplified analysis

libraries:
```{r}
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(pheatmap))
```

Importing data:
```{r}
#set path to file
wd<- "//Cifs2/rcdata$/UKRNASeq/"
countPath<- paste(wd, "RawCounts/CountData.csv", sep = "")

#set sample groups
sampleGroup<- c("pos", "pos", "neg", "pos", "neg", "neg", "neg", "pos", "neg", "neg", "neg", "neg", "neg", "pos", "neg", "neg", rep("control", 5))

#read in count data, format, get rid of NA data, get rid of any genes where all counts are 0
countsInput<- read.csv(countPath)
rownames(countsInput)<- countsInput$X
countsInput<- countsInput[,-1]
colnames(countsInput)<- gsub(".counts.*", "", colnames(countsInput))
counts<- na.omit(countsInput)
counts<- counts[rowSums(counts)>0,]

#create metadata
sample<- colnames(countsInput)
meta<- data.frame(sample, sampleGroup)
rownames(meta)<- meta$sample
```

Creating DESeq2 object and running DESeq2
```{r}
#check that counts/meta data has samples in same order
all(rownames(meta) %in% colnames(counts))
all(rownames(meta) == colnames(counts))

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=counts, 
                             colData=meta, 
                             design=~sampleGroup)

#define levels- so that controls will always be used as reference group
dds$sampleGroup<- relevel(dds$sampleGroup, ref= "control")

#run DE
dds<- DESeq(dds)

#normalize dds for later use. Only need to run this once.
#rld<- rlog(dds)
#write.csv(assay(rld), paste(wd, "rldNorm_AfterDESeq2.csv", sep=''))
```


#Assess data quality
```{r}
#PCA
vsd <- vst(dds, blind=FALSE)
PCAData<-plotPCA(vsd, intgroup="sampleGroup", returnData=TRUE)
#sample 3: PC1 = 3.95 PC2= -1.07
plotPCA(vsd, intgroup="sampleGroup")
```

Get results for all 3 groups
```{r}
#function that takes dds object and desired contrasts as input, outputs ordered results
ContrastDGE<- function(contrasts, dds){
  ContrastsUse<- c("sampleGroup", contrasts)
  ContrastsResults<- na.omit(results(dds, contrast=ContrastsUse))
  resOrdered<- ContrastsResults[order(ContrastsResults$padj),]
  return(resOrdered)
}

#Extract comparisons for 3 desired groups
DGERes_controlpos<- ContrastDGE(c("control", "pos"), dds)
DGERes_controlneg<- ContrastDGE(c("control", "neg"), dds)
DGERes_posneg<- ContrastDGE(c("pos", "neg"), dds)

nrow(DGERes_controlpos[DGERes_controlpos$padj<0.05,])
nrow(DGERes_controlneg[DGERes_controlneg$padj<0.05,])
nrow(DGERes_posneg[DGERes_posneg$padj<0.05,])


```


#Testing different filtering methods, looking at gene dispersion#
This section for testing only, not used for downstream purposes
```{r}
#Test: taking top 75 most significant for posvscontrol/negvscontrol and all of posvsneg-->get gene list --> look at raw counts for these genes

top75_controlpos<- as.data.frame(DGERes_controlpos[1:75,])
top75_controlneg<- as.data.frame(DGERes_controlneg[1:75,])
top75_posneg<- as.data.frame(DGERes_posneg[1:50,])
top75<- unique(c(rownames(top75_controlpos), rownames(top75_controlpos)))
top75<- unique(c(top75, rownames(top75_posneg)))

top75Counts<- counts[top75,]
top75RowSums<- rowSums(top75Counts)
mean(top75RowSums)

#Get raw counts for top 75 genes, split into pos/neg/control. Look at average/SD/dispersion of counts for each gene from within groups
topPos<- top75Counts[,meta$sampleGroup=="pos"]
topNeg<- top75Counts[,meta$sampleGroup=="neg"]
topControl<- top75Counts[,meta$sampleGroup=="control"]

mean(apply(topPos, 1, sd))
#564
mean(apply(topControl, 1, sd))
#1019
mean(apply(topNeg, 1, sd))
#276

posVsControl<- rowMeans(topControl)-rowMeans(topPos)
mean(abs(posVsControl))
#1185

negVsControl<- rowMeans(topControl)-rowMeans(topNeg)
mean(abs(negVsControl))
#1223
mean(apply(topControl, 1, sd))

PosvsNeg<- rowMeans(topNeg)-rowMeans(topPos)
mean(abs(PosvsNeg))
#183



#repeating for all genes
topPos<- countsInput[,meta$sampleGroup=="pos"]
topNeg<- countsInput[,meta$sampleGroup=="neg"]
topControl<- countsInput[,meta$sampleGroup=="control"]

posVsControl<- rowMeans(topControl)-rowMeans(topPos)
mean(abs(posVsControl))
#1185
#all: 117

negVsControl<- rowMeans(topControl)-rowMeans(topNeg)
mean(abs(negVsControl))
#1223
#all: 153

PosvsNeg<- rowMeans(topNeg)-rowMeans(topPos)
mean(abs(PosvsNeg))
#183
#all: 83

mean(abs(rowMeans(topControl)-rowMeans(topPos)))

#comparing to all counts- remove ambiguous/no_feature
allCounts<- counts[1:(nrow(counts)-2),]
mean(rowSums(allCounts))

#average counts for all DEGs
pSig<- DGERes_controlpos[DGERes_controlpos$padj<0.05,]
nSig<- DGERes_controlneg[DGERes_controlneg$padj<0.05,]
pnSig<- DGERes_posneg[DGERes_posneg$padj<0.05,]


allDEGsCounts<- unique(c(rownames(pSig), rownames(nSig)))
allDEGsCounts<- unique(c(allDEGsCounts, rownames(pnSig)))
allDEGs_Counts<-counts[allDEGsCounts,]
mean(rowSums(allDEGs_Counts))

#See summary excel file for mean/min for these results ("FilteringCountsBasedONHeatMapResults_091019")

min(rowSums(top75Counts))
#120

########################################################
#Repeat DESeq using filtered counts data: rowSums > 120#
########################################################

countsInput<- read.csv(countPath)
rownames(countsInput)<- countsInput$X
countsInput<- countsInput[,-1]
colnames(countsInput)<- gsub(".counts.*", "", colnames(countsInput))
counts<- na.omit(countsInput)
counts<- counts[rowSums(counts)>120,]
#15,131 genes

#check that counts/meta data has samples in same order
all(rownames(meta) %in% colnames(counts))
all(rownames(meta) == colnames(counts))

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=counts, 
                             colData=meta, 
                             design=~sampleGroup)

#define levels- so that controls will always be used as reference group
dds$sampleGroup<- relevel(dds$sampleGroup, ref= "control")

#run DE
dds<- DESeq(dds)


#PCA#
vsd <- vst(dds, blind=FALSE)
PCAData<-plotPCA(vsd, intgroup="sampleGroup", returnData=TRUE)
plotPCA(vsd, intgroup="sampleGroup")


#Extract comparisons for 3 desired groups#
DGERes_controlpos<- ContrastDGE(c("control", "pos"), dds)
DGERes_controlneg<- ContrastDGE(c("control", "neg"), dds)
DGERes_posneg<- ContrastDGE(c("pos", "neg"), dds)

nrow(DGERes_controlpos[DGERes_controlpos$padj<0.05,])
nrow(DGERes_controlneg[DGERes_controlneg$padj<0.05,])
nrow(DGERes_posneg[DGERes_posneg$padj<0.05,])

##################
#plotting heatmap#
##################

subtractFromMin<-0
subtractFromMax<- 2.5
lengthOut<- 100
posSig<- DGERes_controlpos[DGERes_controlpos$padj<= 0.05,]
negSig<- DGERes_controlneg[DGERes_controlneg$padj<= 0.05,]
posnegSig<- DGERes_posneg[DGERes_posneg$padj<= 0.05,]

#use top 75 genes in each
topgenes<- unique(c(rownames(negSig[1:75,]), rownames(posSig[1:75,])))
topgenes<- unique(c(topgenes, rownames(posnegSig)))

#rld<- rlog(dds)
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<-0
subtractFromMax<- 2.5
lengthOut<- 100
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```

```{r}
#test for me- if you run DESeq just setting controls/CM, how do samples cluster? Still clustering into pos/neg at all?

#set sample groups
sampleGroup<- c(rep("CM",16), rep("control", 5))

#read in count data, format, get rid of NA data, get rid of any genes where all counts are 0
countsInput<- read.csv(countPath)
rownames(countsInput)<- countsInput$X
countsInput<- countsInput[,-1]
colnames(countsInput)<- gsub(".counts.*", "", colnames(countsInput))
counts<- na.omit(countsInput)
counts<- counts[rowSums(counts)>0,]

#create metadata
sample<- colnames(countsInput)
meta<- data.frame(sample, sampleGroup)
rownames(meta)<- meta$sample

#check that counts/meta data has samples in same order
all(rownames(meta) %in% colnames(counts))
all(rownames(meta) == colnames(counts))

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=counts, 
                             colData=meta, 
                             design=~sampleGroup)

#define levels- so that controls will always be used as reference group
dds$sampleGroup<- relevel(dds$sampleGroup, ref= "control")

#run DE
dds<- DESeq(dds)


#PCA#
vsd <- vst(dds, blind=FALSE)
PCAData<-plotPCA(vsd, intgroup="sampleGroup", returnData=TRUE)
plotPCA(vsd, intgroup="sampleGroup")

#heatmap#
res<- as.data.frame(na.omit(results(dds))) 
res<- res[order(res$padj),]
topSig<- res[1:150,]


#rld<- rlog(dds)
rld<- rlog(dds)
mat<- assay(rld)[rownames(topSig),]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))

```

```{r}
#Test: using genes with rowsums of raw counts > 0 MINUS genes that are not in fpkm data

#set sample groups
sampleGroup<- c("pos", "pos", "neg", "pos", "neg", "neg", "neg", "pos", "neg", "neg", "neg", "neg", "neg", "pos", "neg", "neg", rep("control", 5))

#read in count data, format
countsInput<- read.csv(countPath)
rownames(countsInput)<- countsInput$X
countsInput<- countsInput[,-1]
colnames(countsInput)<- gsub(".counts.*", "", colnames(countsInput))
dim(countsInput)
#26,365

#get rid of any genes from counts that aren't in fpkm
fpkm<- read.csv(paste(wd, "fpkm_tracking/fpkmData.csv", sep =""))
dim(fpkm)
#26,228
GenesInFPKM_NotinCounts<- rownames(countsInput) %in% fpkm$X
length(GenesInFPKM_NotinCounts)
#26,365
sum(GenesInFPKM_NotinCounts)
#26,227- 26,227 genes in counts are also in fpkm
#there is one gene that's in fpkm but is not in counts- DDX11L1. It would have been excluded from the previous analysis that used fpkm filtering, and it's excluded here.
GenesInFPKM_NotinCounts<- countsInput[GenesInFPKM_NotinCounts,]
#gets rid of 138 genes


rownames(fpkm)<- fpkm$X
fpkm<- fpkm[,-1]
fpkmFilter<- fpkm[rowMeans(fpkm)>0,]
fpkmFilter<- rownames(countsInput)%in% rownames(fpkmFilter)
GenesInFPKM_NotinCounts<-countsInput[fpkmFilter,]

counts<- GenesInFPKM_NotinCounts[rowSums(GenesInFPKM_NotinCounts)>0,]
#22571 genes vs 22575 if no fpkm filtering done

counts_lessthan0<- countsInput[rowSums(countsInput)>0,]
test<-rownames(counts_lessthan0) %in% rownames(counts)
fourgenes<- counts_lessthan0[!test,]
#four genes that are omitted when you do this type of filtering: no_feature, ambiguous, MIR548C, MIR447A- both miRNAs


#create metadata
sample<- colnames(countsInput)
meta<- data.frame(sample, sampleGroup)
rownames(meta)<- meta$sample

#check that counts/meta data has samples in same order
all(rownames(meta) %in% colnames(counts))
all(rownames(meta) == colnames(counts))

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=counts, 
                             colData=meta, 
                             design=~sampleGroup)

#define levels- so that controls will always be used as reference group
dds$sampleGroup<- relevel(dds$sampleGroup, ref= "control")

#run DE
dds<- DESeq(dds)

#get results
DGERes_controlpos<- ContrastDGE(c("control", "pos"), dds)
DGERes_controlneg<- ContrastDGE(c("control", "neg"), dds)
DGERes_posneg<- ContrastDGE(c("pos", "neg"), dds)

nrow(DGERes_controlpos[DGERes_controlpos$padj<0.05,])
nrow(DGERes_controlneg[DGERes_controlneg$padj<0.05,])
nrow(DGERes_posneg[DGERes_posneg$padj<0.05,])

#heatmap

posSig<- DGERes_controlpos[DGERes_controlpos$padj<= 0.05,]
negSig<- DGERes_controlneg[DGERes_controlneg$padj<= 0.05,]
posnegSig<- DGERes_posneg[DGERes_posneg$padj<= 0.05,]

rld<- rlog(dds)
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```

Write to csv
```{r}
write.csv(DGERes_controlpos, paste(wd, "DESeq2/Final/", "DGERes_controlvspos.csv", sep = ""))
write.csv(DGERes_controlneg, paste(wd, "DESeq2/Final/", "DGERes_controlvsneg.csv", sep = ""))
write.csv(DGERes_posneg, paste(wd, "DESeq2/Final/", "DGERes_posvsneg.csv", sep = ""))
```


#Clustering DESeq2 results in heatmap

Select only significant DEGs
```{r}
posSig<- DGERes_controlpos[DGERes_controlpos$padj<= 0.05,]
negSig<- DGERes_controlneg[DGERes_controlneg$padj<= 0.05,]
posnegSig<- DGERes_posneg[DGERes_posneg$padj<= 0.05,]
```

Select only DEGs with logfoldchange above 1
```{r}
Above1Pos<- posSig[abs(posSig$log2FoldChange)>=1,]
Above1Neg<- negSig[abs(negSig$log2FoldChange)>=1,]
Above1PosNeg<- posnegSig[abs(posnegSig$log2FoldChange)>=1,]

subtractFromMin<-0
subtractFromMax<- 2.5
lengthOut<- 75
```


Plotting heatmaps: multiple options

#Option #1: Using all sig genes
```{r}
topgenes<- unique(c(rownames(negSig), rownames(posSig)))
topgenes<- unique(c(topgenes, rownames(posnegSig)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,9
```


#Option #2: Using top 50 sig genes from each
```{r}
topgenes<- unique(c(rownames(negSig[1:50,]), rownames(posSig[1:50,])))
topgenes<- unique(c(topgenes, rownames(posnegSig[1:50,])))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,3,5,9,17
```



#Option #3: Using top 100 sig genes from pos/neg vs control and all of pos vs neg
```{r}
topgenes<- unique(c(rownames(negSig[1:100,]), rownames(posSig[1:100,])))
topgenes<- unique(c(topgenes, rownames(posnegSig)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,3,5,9,17
```
#Option #4: Using top 75 sig genes from pos/neg vs control and all of pos vs neg
```{r}
topgenes<- unique(c(rownames(negSig[1:75,]), rownames(posSig[1:75,])))
topgenes<- unique(c(topgenes, rownames(posnegSig)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,3,5,9,17
```

#Option #4: Using top 200 sig genes from pos/neg vs control and all of pos vs neg
```{r}
topgenes<- unique(c(rownames(negSig[1:200,]), rownames(posSig[1:200,])))
topgenes<- unique(c(topgenes, rownames(posnegSig)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,9
```

#Option #4: Using top 150 sig genes from pos/neg vs control and all of pos vs neg
```{r}
topgenes<- unique(c(rownames(negSig[1:150,]), rownames(posSig[1:150,])))
topgenes<- unique(c(topgenes, rownames(posnegSig)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,9
```

#Option #4: Using top 125 sig genes from pos/neg vs control and all of pos vs neg
```{r}
topgenes<- unique(c(rownames(negSig[1:125,]), rownames(posSig[1:125,])))
topgenes<- unique(c(topgenes, rownames(posnegSig)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: control / 1,2,5,9,3,17
```

#Option #4: Use sig genes with fold change above 1
```{r}
topgenes<- unique(c(rownames(Above1Neg), rownames(Above1Pos)))
topgenes<- unique(c(topgenes, rownames(Above1PosNeg)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: controls/1,2,9
```

#Option #5: Use sig genes with fold change above .5
```{r}
Above1Pos<- posSig[abs(posSig$log2FoldChange)>=.5,]
Above1Neg<- negSig[abs(negSig$log2FoldChange)>=.5,]
Above1PosNeg<- posnegSig[abs(posnegSig$log2FoldChange)>=.5,]

topgenes<- unique(c(rownames(Above1Neg), rownames(Above1Pos)))
topgenes<- unique(c(topgenes, rownames(Above1PosNeg)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: controls/1,2,9
```


#Option #5: Use sig genes with fold change above .1
```{r}
Above1Pos<- posSig[abs(posSig$log2FoldChange)>=.1,]
Above1Neg<- negSig[abs(negSig$log2FoldChange)>=.1,]
Above1PosNeg<- posnegSig[abs(posnegSig$log2FoldChange)>=.1,]

topgenes<- unique(c(rownames(Above1Neg), rownames(Above1Pos)))
topgenes<- unique(c(topgenes, rownames(Above1PosNeg)))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
#clusters: controls/1,2,9
```


Conclusion: using a log2 fold change cutoff isn't super useful; doesn't seem to improve clustering.





