---
title: "UKy miRNA: DESeq2 and RNASeq integration"
output: html_notebook
---

#Running DESeq2 on miRNA data + using multiMiR to integrate with RNASeq data

Here, the following input is used for DESeq2:

isomirs --> remove seqs --> collapse together, adding counts together 

+                                                                              

mature 


Other filtering/combinations were tried (see miRNA_DESeq2_HitsIsomirsFamilies notebook), but this method was simple, gave good clustering, and treats all species the same, so we decided to use this method moving forward. Repeating here for the sake of completeness.

#Libraries
```{r}
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(pheatmap))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(multiMiR))
```


#Running DESeq2: first, filter/combine using the method described above


Read in data- one file here, have already been combined into one after downloading separately for each sample
```{r}
wd<- "//Cifs2/rcdata$/UKRNASeq/miRNA/RawData/"
MatureHits<- read.csv(paste(wd, "MaturemiRNA/miRNA_matureHits_downloadedSeparately.csv", sep =""))
rownames(MatureHits)<- MatureHits$X
MatureHits<- MatureHits[,-1]
colnames(MatureHits)<- c(colnames(MatureHits)[1:23], "control.5")

IsomirHits<- read.csv(paste(wd, "IsoHits/miRNA_isomirs_downloadedSeparately.csv", sep =""))
rownames(IsomirHits)<- IsomirHits$X
IsomirHits<- IsomirHits[,-1]
```


Filter data (collapsed isomirs added to whole list of mature)
```{r}
##################
#Collapse isomirs#
##################
hits<- IsomirHits

#get rid of any IDs that have 0 counts for all samples
hits<- hits[rowSums(hits)>0,]
hitsFinal<- hits

hitsFinal$miRNA<- gsub(".*_", "",rownames(hitsFinal))

df<- as_tibble(hitsFinal)
df<-df%>% group_by(miRNA) %>% summarise_at(colnames(hits)[1:24],sum, na.rm=TRUE)
hits<- as.data.frame(df)
rownames(hits)<-hits$miRNA
hits<- hits[,-1]


#####################
#Combine with mature#
#####################

#rbind with Mature hits
FinalHits<- rbind(hits,MatureHits)

#Remove sample 17 (very low counts)
FinalHits<- FinalHits[,!colnames(FinalHits)%in% c("CM10", "CM14", "CM4", "CM17")]
```

Run DESeq2 using Final Hits as input
```{r}
hits<-FinalHits

#set sample groups
sampleGroup<- c(rep("CM", 15), rep("control", 5))

#create metadata
sample<- colnames(hits)
meta<- data.frame(sample, sampleGroup)
rownames(meta)<- meta$sample

#Run DESeq2, extract results
dds<- DESeqDataSetFromMatrix(countData=hits, 
                             colData=meta, 
                             design=~sampleGroup)

dds$sampleGroup<- relevel(dds$sampleGroup, ref= "control")

dds<- DESeq(dds)

res<- as.data.frame(results(dds))
res<- na.omit(res)
res<- res[order(res$padj),]

nrow(res[res$padj<0.05,])
#11

#To use for multmiR
miRNA_DE<- res

summary(results(dds))
```

Cluster samples using PCA
```{r}
#Function that takes hit counts and metadata as input, outputs PCA plot
myPCA<- function(hits, meta){
  if (all(rownames(meta) %in% colnames(hits))){
    if (all(rownames(meta) == colnames(hits))){
      dds<- DESeqDataSetFromMatrix(countData=hits, 
                             colData=meta, 
                             design=~sampleGroup)
      dds$sampleGroup<- relevel(dds$sampleGroup, ref= "control")
      rld<- rlog(dds)
      plotPCA(rld, intgroup="sampleGroup")
    }
  } else {
    print ("Error: check metadata")
  }
}

myPCA(hits,meta)
```

```{r}
#heatmap
rld<- rlog(dds)
topgenes<- rownames(res[1:25,])
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<- -.5
subtractFromMax<- 2.5
lengthOut<- 100
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```


#Running DESeq2 with all 3 groups (TTN+/-) instead of just two
```{r}
#set sample groups
sampleGroup<- c("pos", "pos", "neg", "pos", "neg", "neg", "neg", "pos", "neg", "neg", "neg", "neg", "neg", "neg", "neg", rep("control", 5))

#create metadata
sample<- colnames(hits)
meta<- data.frame(sample, sampleGroup)
rownames(meta)<- meta$sample

#Run DESeq2, extract results
dds<- DESeqDataSetFromMatrix(countData=hits, 
                             colData=meta, 
                             design=~sampleGroup)

dds$sampleGroup<- relevel(dds$sampleGroup, ref= "control")

dds<- DESeq(dds)

PosvControl<- results(dds, contrast = c("sampleGroup", "pos", "control"))
PosvControl<- na.omit(PosvControl[order(PosvControl$padj),])
nrow(PosvControl[PosvControl$padj<0.05,])
#0

NegvControl<- results(dds, contrast = c("sampleGroup", "neg", "control"))
NegvControl<- na.omit(NegvControl[order(NegvControl$padj),])
nrow(NegvControl[NegvControl$padj<0.05,])
#17

PosvNeg<- results(dds, contrast = c("sampleGroup", "pos", "neg"))
PosvNeg<- na.omit(PosvNeg[order(PosvNeg$padj),])
nrow(PosvNeg[PosvNeg$padj<0.05,])
#1
```
```{r}
myPCA(hits,meta)
```

```{r}
#heatmap
rld<- rlog(dds)
TopGenesUse<- 10
TopGenes<- unique(c(rownames(PosvControl)[1:TopGenesUse],rownames(PosvNeg)[1:TopGenesUse],rownames(NegvControl)[1:TopGenesUse]))

mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

subtractFromMin<- -.5
subtractFromMax<- 2.5
lengthOut<- 100
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)))
```


Write final list of miRNAs to file (sorted by padj, but including all species):
```{r}
write.csv(res, "//Cifs2/rcdata$/UKRNASeq/miRNA/DESeq2/UKy_miRNADESeq2Results_MatureCollapsedIsomirs_100219.csv")
```



#Integrating with RNASeq data

First: run DESeq2 on mRNA data, using just 2 groups (3 used previously, makes more sense to use 2 here):

```{r}
#set sample groups
sampleGroup<- c(rep("CM", 16), rep("control", 5))

#read in count data, format, get rid of NA data, get rid of any genes where all counts are 0
countsInput<- read.csv("//Cifs2/rcdata$/UKRNASeq/RawCounts/CountData.csv")
rownames(countsInput)<- countsInput$X
countsInput<- countsInput[,-1]
colnames(countsInput)<- gsub(".counts.*", "", colnames(countsInput))
counts<- na.omit(countsInput)
counts<- counts[rowSums(counts)>0,]

#create metadata
sample<- colnames(countsInput)
meta<- data.frame(sample, sampleGroup)
rownames(meta)<- meta$sample

#check that counts/meta data has samples in same order
all(rownames(meta) %in% colnames(counts))
all(rownames(meta) == colnames(counts))

#create DESeq2 object
dds<- DESeqDataSetFromMatrix(countData=counts, 
                             colData=meta, 
                             design=~sampleGroup)

#define levels- so that controls will always be used as reference group
dds$sampleGroup<- relevel(dds$sampleGroup, ref= "control")

#run DE
dds<- DESeq(dds)

res<- as.data.frame(results(dds))
res<- res[order(res$padj),]
nrow(res[res$padj<0.05,])
#9,000 DE

#write DEGs to csv
write.csv(res, "//Cifs2/rcdata$/UKRNASeq/DESeq2/Final/DGERes_ControlvsCm.csv")
```

Read in miRNA and mRNA 2 group DESeq2 results- don't need to run the previous chunk in the future
```{r}
mRNADEGs<- read.csv("//Cifs2/rcdata$/UKRNASeq/DESeq2/Final/DGERes_ControlvsCm.csv")
rownames(mRNADEGs)<- mRNADEGs$X
mRNADEGs<- mRNADEGs[,-1]

miRNADEGs<-read.csv("//Cifs2/rcdata$/UKRNASeq/miRNA/DESeq2/miRNADESeq2Results_matureInput.csv")
rownames(miRNADEGs)<- miRNADEGs$X
miRNADEGs$X<- tolower(miRNADEGs$X)

#select just significant mDEGs
sigMDEGs<- na.omit(mRNADEGs[mRNADEGs$padj<0.05,])

#Select just DE miRNAs
sigMiDEGs<- miRNADEGs[miRNADEGs$padj<0.05,]
```

Running multiMiR
```{r}
#run multiMiR
multimir_1<- get_multimir(mirna=sigMiDEGs$X, summary=TRUE)

head(multimir_1@data)
```
```{r}
data<- multimir_1@data

#how many miRNAs have at least 1 target in this list?
length(unique(data$mature_mirna_id))
#9- 1 is isomir

#how many targets are there?
length(data$target_ensembl)
#1210

#Get list of targets. How many of these targets are in the list of mRNA DEGs?
TargetsinDEGs<- data$target_symbol %in% rownames(sigMDEGs)
sum(TargetsinDEGs)
#288

#If the DE miRNAs are really downregulating the mRNAs, then you would expect that miRNAs that are UPREGULATED in CM would have targets that are DOWNREGULATED in CM
upmiRNAs<- sigMiDEGs[sigMiDEGs$log2FoldChange>0,]
nrow(upmiRNAs)
#6 miRNAs that are upregulated in CM

upmiRNATargets<- data[tolower(data$mature_mirna_id) %in% upmiRNAs$X,]
nrow(upmiRNATargets)
#1018 targets of these upregulated miRNAs

unique(upmiRNATargets$mature_mirna_id)
#6 of these upregulated have at least one target

downmRNAs<- sigMDEGs[sigMDEGs$log2FoldChange<0,]
nrow(downmRNAs)
#2,240 downregulated mRNAs in CM

#how many of the targets of up miRNAs are in down mRNAs?
UpTargetsinDownDEGs<- upmiRNATargets$target_symbol %in% rownames(downmRNAs)
sum(UpTargetsinDownDEGs)
#124

#targets have different ratings of strength. Of the targets from miRNAs that are upregulated that are downregulated in mRNA, which have a higher confidence?
downmRNAs<- upmiRNATargets[UpTargetsinDownDEGs,]
weak<- grepl("Weak", downmRNAs$support_type)
notWeak<- downmRNAs[!weak,]
length(unique(notWeak$target_symbol))
#16 such targets
notWeak$target_symbol
```

```{r}
#Doing the opposite: miRNAs that are significantly DOWNREGULATED in CM --> their targets --> matches in significantly UPREGULATED mRNAs
downmiRNAs<- sigMiDEGs[sigMiDEGs$log2FoldChange<0,]
nrow(downmiRNAs)
#2 miRNAs that are downregulated in CM

downmiRNATargets<- data[data$mature_mirna_id%in% rownames(downmiRNAs),]
nrow(downmiRNATargets)
#191 targets of these upregulated miRNAs

unique(downmiRNATargets$mature_mirna_id)
#2 of these 2 downregulated miRNAs have targets in the list

upmRNAs<- sigMDEGs[sigMDEGs$log2FoldChange>0,]
nrow(upmRNAs)
#1,925 upregulated mRNAs in CM

#how many of the targets of up miRNAs are in down mRNAs?
DownTargetsinupDEGs<- downmiRNATargets$target_symbol %in% rownames(upmRNAs)
sum(DownTargetsinupDEGs)
#25

#targets have different ratings of strength. Of the targets from miRNAs that are upregulated that are downregulated in mRNA, which have a higher confidence?
upmRNAs<- downmiRNATargets[DownTargetsinupDEGs,]
weak<- grepl("Weak", upmRNAs$support_type)
notWeak<- upmRNAs[!weak,]
length(unique(notWeak$target_symbol))
#5 such targets
unique(notWeak$target_symbol)

```

```{r}
#Instead of using the confidence in the assay to determine the final list of affected mRNA targets, look at the targets that's affected by MORE THAN ONE miRNA

#for upregulated miRNA targets in downregulated mRNAs:
NumberOfDownOccurences<-table(downmRNAs$target_symbol, downmRNAs$mature_mirna_id)
NumberofupmiRNAs<-apply(NumberOfDownOccurences,1,function(x) sum(x > 0))
sum(NumberofupmiRNAs>1)
#1 mRNA has more than potential miRNA match
NumberofupmiRNAs[NumberofupmiRNAs>1]
#ABHD15 

#for downregulated miRNA targets in upregulated mRNAs:
NumberOfUpOccurences<-table(upmRNAs$target_symbol, upmRNAs$mature_mirna_id)
NumberofdownmiRNAs<-apply(NumberOfUpOccurences,1,function(x) sum(x > 0))
sum(NumberofdownmiRNAs>1)
#0 mRNAs has more than potential miRNA match

```

```{r}
#write results to csv- combining up regulated miRNAs with their downregulated mRNA targets and downregulated miRNAs with their upregualted mRNA targets

upmiRNADownmRNA<- downmRNAs[order(downmRNAs$mature_mirna_id), c(2,3,4,6,7,8)]
upmiRNADownmRNA$miRNALog2<- "Positive"
upmiRNADownmRNA$mRNALog2<- "Negative"

downmiRNAUpmRNA<- upmRNAs[order(upmRNAs$mature_mirna_id), c(2,3,4,6,7,8)]
downmiRNAUpmRNA$miRNALog2<- "Negative"
downmiRNAUpmRNA$mRNALog2<- "Positive"

write.csv(rbind(upmiRNADownmRNA, downmiRNAUpmRNA), "//Cifs2/rcdata$/UKRNASeq/miRNA/multiMiR/multimirResults_MatureInput.csv" )
```


#Running same analysis, but with miRNA list from using just mature miRNA as input for DESeq2

```{r}
miRNADEGs<-read.csv("//Cifs2/rcdata$/UKRNASeq/miRNA/DESeq2/miRNADESeq2Results_matureInput.csv")
rownames(miRNADEGs)<- miRNADEGs$X
miRNADEGs$X<- tolower(miRNADEGs$X)

#select just significant mDEGs
sigMDEGs<- na.omit(mRNADEGs[mRNADEGs$padj<0.05,])

#Select just DE miRNAs
sigMiDEGs<- miRNADEGs[miRNADEGs$padj<0.05,]
```

Running multiMiR
```{r}
#run multiMiR
multimir_1<- get_multimir(mirna=sigMiDEGs$X, summary=TRUE)

head(multimir_1@data)
```
```{r}
data<- multimir_1@data

#how many miRNAs have at least 1 target in this list?
length(unique(data$mature_mirna_id))
#9- 1 more than the number of DE miRNA bc targets were found for both the mature and the isomir for hsa-mir-126-3p

#how many targets are there?
length(data$target_ensembl)
#1210

#Get list of targets. How many of these targets are in the list of mRNA DEGs?
TargetsinDEGs<- data$target_symbol %in% rownames(sigMDEGs)
sum(TargetsinDEGs)
#288

#If the DE miRNAs are really downregulating the mRNAs, then you would expect that miRNAs that are UPREGULATED in CM would have targets that are DOWNREGULATED in CM
upmiRNAs<- sigMiDEGs[sigMiDEGs$log2FoldChange>0,]
nrow(upmiRNAs)
#6 miRNAs that are upregulated in CM

upmiRNATargets<- data[tolower(data$mature_mirna_id) %in% upmiRNAs$X,]
nrow(upmiRNATargets)
#1018 targets of these upregulated miRNAs

unique(upmiRNATargets$mature_mirna_id)
#6 of these upregulated have at least one target

downmRNAs<- sigMDEGs[sigMDEGs$log2FoldChange<0,]
nrow(downmRNAs)
#2,240 downregulated mRNAs in CM

#how many of the targets of up miRNAs are in down mRNAs?
UpTargetsinDownDEGs<- upmiRNATargets$target_symbol %in% rownames(downmRNAs)
sum(UpTargetsinDownDEGs)
#124

#targets have different ratings of strength. Of the targets from miRNAs that are upregulated that are downregulated in mRNA, which have a higher confidence?
downmRNAs<- upmiRNATargets[UpTargetsinDownDEGs,]
weak<- grepl("Weak", downmRNAs$support_type)
notWeak<- downmRNAs[!weak,]
nrow(notWeak)
#18 such targets
unique(notWeak$target_symbol)
# [1] "ZWINT"  "SCRIB"  "SCRIB"  "NFYB"   "TYMS"   "CFTR"   "HIF1A"  "NCAM1"  "INPP5A" "LDHA"   "PTPN1"  "TFRC"  
#[13] "PTBP3"  "GPD1L"  "FGFRL1" "FGFRL1" "P4HB"   "KCMF1"
```

```{r}
#Doing the opposite: miRNAs that are significantly DOWNREGULATED in CM --> their targets --> matches in significantly UPREGULATED mRNAs
downmiRNAs<- sigMiDEGs[sigMiDEGs$log2FoldChange<0,]
nrow(downmiRNAs)
#2 miRNAs that are downregulated in CM

downmiRNATargets<- data[data$mature_mirna_id%in% rownames(downmiRNAs),]
nrow(downmiRNATargets)
#191 targets of these downregulated miRNAs

unique(downmiRNATargets$mature_mirna_id)
#both of the 2 miRNAs that are downregulated have targets

upmRNAs<- sigMDEGs[sigMDEGs$log2FoldChange>0,]
nrow(upmRNAs)
#1,925 upregulated mRNAs in CM

#how many of the targets of up miRNAs are in down mRNAs?
DownTargetsinupDEGs<- downmiRNATargets$target_symbol %in% rownames(upmRNAs)
sum(DownTargetsinupDEGs)
#25

#targets have different ratings of strength. Of the targets from miRNAs that are upregulated that are downregulated in mRNA, which have a higher confidence?
upmRNAs<- downmiRNATargets[DownTargetsinupDEGs,]
weak<- grepl("Weak", upmRNAs$support_type)
notWeak<- upmRNAs[!weak,]
nrow(notWeak)
#18 such targets
unique(notWeak$target_symbol)
#"IRS1"   "VEGFA"  "CXCL12" "LRP6"   "SOX6"

```
```{r}
#Combine up regulated miRNAs with their downregulated mRNA targets and downregulated miRNAs with their upregualted mRNA targets

upmiRNADownmRNA<- downmRNAs[order(downmRNAs$mature_mirna_id), c(2,3,4,6,7,8)]
upmiRNADownmRNA$miRNALog2<- "Positive"
upmiRNADownmRNA$mRNALog2<- "Negative"

downmiRNAUpmRNA<- upmRNAs[order(upmRNAs$mature_mirna_id), c(2,3,4,6,7,8)]
downmiRNAUpmRNA$miRNALog2<- "Negative"
downmiRNAUpmRNA$mRNALog2<- "Positive"

miRNATargets_mRNASig<- rbind(upmiRNADownmRNA, downmiRNAUpmRNA)


#add in 4 additional columns- log2 and padj for mRNAs in list, log2 and padj for miRNAs in list
#First: get just the log2fc and padj values for sig degs with target matches (i.e., all target in the "mirnaTargets_mrnaSig" file)
mrna_log2fc_padj<- sigMDEGs[rownames(sigMDEGs) %in% miRNATargets_mRNASig$target_symbol, c(2,6)]
colnames(mrna_log2fc_padj)<- c("mRNA_log2FC", "mRNA_padj")
mrna_log2fc_padj$target_symbol<- rownames(mrna_log2fc_padj)

#merge this with the mirnatargets_mrnasig file- multimir data with log2/padj data
miRNATargets_mRNASig_mrnalogpadj<- merge(miRNATargets_mRNASig, mrna_log2fc_padj, by= "target_symbol")

#repeat for the miRNA values
mirna_log2fc_padj<- sigMiDEGs[rownames(sigMiDEGs) %in% miRNATargets_mRNASig$mature_mirna_id, c(3,7)]
colnames(mirna_log2fc_padj)<- c("miRNA_log2FC", "miRNA_padj")
mirna_log2fc_padj$mature_mirna_id<- rownames(mirna_log2fc_padj)

#merge this with the mirnatargets_mrnasig file- multimir data with log2/padj data
miRNATargets_mRNASig_mrnalogpadj_mirnalogpadj<- merge(miRNATargets_mRNASig_mrnalogpadj, mirna_log2fc_padj, by= "mature_mirna_id")


```


```{r}
write.csv(miRNATargets_mRNASig_mrnalogpadj_mirnalogpadj, "//Cifs2/rcdata$/UKRNASeq/miRNA/multiMiR/multimirResults_MatureInput_log2padj.csv" )
```

```{r}
BiocManager::install("org.Hs.eg.db")
library("org.Hs.eg.db")

ncbiGeneID<- mapIds(org.Hs.eg.db, "SOX6", "ENTREZID", "SYMBOL")

BiocManager::install("KEGGREST")
library(KEGGREST)

listDatabases()

query<- keggGet("hsa:55553")
```

```{r}

```

