---
title: "FINAL: Integrating RNASeq and miRNASeq DE Results- UKy"
output: html_notebook
---

#Combining DESeq2 results from RNASeq and miRNASeq analyses

Comparing miRNA targets to mRNA DEGs

**Import libraries**
```{r}
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(pheatmap))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(multiMiR))
```


**Import DESeq2 results for Controls vs CM analysis for RNASeq and miRNASeq**
```{r}
#set wd
wd<- "//Cifs2/rcdata$/UKRNASeq/"

#Read in RNASeq DESeq2 Results (CM vs Controls), format
RNADESeq2Results<- read.csv(paste(wd, "DESeq2_mRNA/DEResults_ControlsvsCM_AllGenes.csv", sep =""))
rownames(RNADESeq2Results)<- as.character(RNADESeq2Results$X)
RNADESeq2Results<- RNADESeq2Results[,c(2:7)]

#Read in miRNASeq DESeq2 Results (CM vs Controls), format
miRNADESeq2Results<- read.csv(paste(wd, "DESeq2_miRNA/miRNADESeq2Results_ControlsvsCM.csv", sep =""))
rownames(miRNADESeq2Results)<- miRNADESeq2Results$X


#select just significantly DE mRNA
sigMDEGs<- na.omit(RNADESeq2Results[RNADESeq2Results$padj<0.05,])

#Select just significantly DE miRNAs
sigMiDEGs<- miRNADESeq2Results[miRNADESeq2Results$padj<0.05,]
```

**Determine mRNA targets of sig DE miRNAs: using multiMiR package**

Searches through 16 databases to find all targets of input miRNAs
```{r}
#run multiMiR
multimirResults<- get_multimir(mirna=sigMiDEGs$X, summary=TRUE)

head(multimirResults@data)
```


**Explore multimir data:**
```{r}
data<- multimirResults@data

length(unique(data$mature_mirna_id))
#Targets for all 19 miRNA species 

upmiRNAs<- sigMiDEGs[sigMiDEGs$log2FoldChange>0,]
nrow(upmiRNAs)
#11 miRNAs up regulated in CM

downmiRNAs<- sigMiDEGs[sigMiDEGs$log2FoldChange<0,]
nrow(downmiRNAs)
#8 miRNAs downregulated in CM

upmiRNATargets<- data[data$mature_mirna_id %in% upmiRNAs$X,]
length(unique(upmiRNATargets$target_symbol))
#3,410 mRNA targets of the 11 upregulated miRNAs

downmiRNATargets<- data[data$mature_mirna_id%in% downmiRNAs$X,]
length(unique(downmiRNATargets$target_symbol))
#6,244 mRNA targets of the 8 downregulated miRNAs

downmRNAs<- sigMDEGs[sigMDEGs$log2FoldChange<0,]
nrow(downmRNAs)
#2,532 mRNAs downregulated in CM

upmRNAs<- sigMDEGs[sigMDEGs$log2FoldChange>0,]
nrow(upmRNAs)
#1,989 mRNAs upregulated in CM

UpTargetsinDownDEGs<- upmiRNATargets$target_symbol %in% rownames(downmRNAs)
sum(UpTargetsinDownDEGs)
#Of the 5,365 UPREGULATED miRNA targets, 775 of them are also DOWNREGULATED in the mRNA data

DownTargetsinupDEGs<- downmiRNATargets$target_symbol %in% rownames(upmRNAs)
sum(DownTargetsinupDEGs)
#Of the 11,327 DOWNREGULATED miRNA targets, 1,339 of them are also UPREGULATED in the mRNA data
```

Important takeaways:
There are 11 upregulated miRNAs with a total of 5,365 mRNA targets between them. Of these targets, 1,989 of them are also downregulated in the mRNA data

There are 8 downregulated miRNAs with a total of 11,327 mRNA targets between them. Of these targets, 1,339 of them are also upregulated in the mRNA data

That gives us 3,328 potentially interesting mRNAs- that is, mRNAs that are significantly different in CM that may be being regulated by miRNAs that are also significantly different in CM

Of those 3,328 mRNAs: which to focus on?

**One way to narrow down: the database that searches for miRNA/target interactions has a ranking of how "strong" the interaction is- some labeled as "weak". Looking at only the "strong":**
```{r}
downTargets<- upmiRNATargets[UpTargetsinDownDEGs,]
weakDownTargets<- grepl("Weak", downTargets$support_type)
notWeakDownTargets<- downTargets[!weakDownTargets,]
length(unique(notWeakDownTargets$target_symbol))
#289 of the 1,989 "interesting" downregulated mRNA targets have a STRONG support type for interacting with their upregulated miRNA

upTargets<- downmiRNATargets[DownTargetsinupDEGs,]
weakUpTargets<- grepl("Weak", upTargets$support_type)
notWeakUpTargets<- upTargets[!weakUpTargets,]
length(unique(notWeakUpTargets$target_symbol))
#619 of the 1,339 "interesting" upregulated mRNA targets have a STRONG support type for interacting with their downregulated miRNA
```



**Prep these results for exporting to a file- need to combine the two sets of results and include info on the padj/log2FC of the miRNA/mRNA species**
```{r}
#Combine up regulated miRNAs with their downregulated mRNA targets and downregulated miRNAs with their upregualted mRNA targets
upmiRNADownmRNA<- downTargets[order(downTargets$mature_mirna_id), c(2,3,4,6,7,8)]
upmiRNADownmRNA$miRNALog2<- "Positive"
upmiRNADownmRNA$mRNALog2<- "Negative"

downmiRNAUpmRNA<- upTargets[order(upTargets$mature_mirna_id), c(2,3,4,6,7,8)]
downmiRNAUpmRNA$miRNALog2<- "Negative"
downmiRNAUpmRNA$mRNALog2<- "Positive"

miRNATargets_mRNASig<- rbind(upmiRNADownmRNA, downmiRNAUpmRNA)


#add in 4 additional columns- log2 and padj for mRNAs in list, log2 and padj for miRNAs in list
#First: get just the log2fc and padj values for sig degs with target matches (i.e., all targets in the "mirnaTargets_mrnaSig" file)
mrna_log2fc_padj<- sigMDEGs[rownames(sigMDEGs) %in% miRNATargets_mRNASig$target_symbol, c(2,6)]
colnames(mrna_log2fc_padj)<- c("mRNA_log2FC", "mRNA_padj")
mrna_log2fc_padj$target_symbol<- rownames(mrna_log2fc_padj)

#merge this with the mirnatargets_mrnasig file- multimir data with log2/padj data
miRNATargets_mRNASig_mrnalogpadj<- merge(miRNATargets_mRNASig, mrna_log2fc_padj, by= "target_symbol")

#repeat for the miRNA values- get padj and log2FC values
mirna_log2fc_padj<- sigMiDEGs[rownames(sigMiDEGs) %in% miRNATargets_mRNASig$mature_mirna_id, c(3,7)]
colnames(mirna_log2fc_padj)<- c("miRNA_log2FC", "miRNA_padj")
mirna_log2fc_padj$mature_mirna_id<- rownames(mirna_log2fc_padj)

#merge this with the mirnatargets_mrnasig file- multimir data with log2/padj data
miRNATargets_mRNASig_mrnalogpadj_mirnalogpadj<- merge(miRNATargets_mRNASig_mrnalogpadj, mirna_log2fc_padj, by= "mature_mirna_id")

```


**Export this file to csv**
```{r}
write.csv(miRNATargets_mRNASig_mrnalogpadj_mirnalogpadj, paste(wd, "DESeq2_miRNA/multiMiR/multimirResults_log2padj.csv", sep =""))
```

```{r}
sessionInfo()
```

