---
title: "R Notebook"
output: html_notebook
---

Import libraries:
```{r}
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(pheatmap))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(metaseqR))
suppressPackageStartupMessages(library(ggVennDiagram))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(fgsea))
suppressPackageStartupMessages(library(dplyr))
```




Functions
```{r}
#Create DESeq2 object, check that metadata is correct, and run DESeq2
RunDESeq2<- function(counts, meta){
  #check that counts/meta data has samples in same order
  print(all(rownames(meta) %in% colnames(counts)))
  print(all(rownames(meta) == colnames(counts)))
  
  #create DESeq2 object
  dds<- DESeqDataSetFromMatrix(countData=counts, 
                               colData=meta, 
                               design=~Sex + sampleGroup)
  
  #define levels- so that controls will always be used as reference group
  dds$sampleGroup<- relevel(dds$sampleGroup, ref= "Control")
  
  #run DE
  dds<- DESeq(dds)
  
  return(dds)

}


#Plot PCA nicely
MyPCA<- function(dds, metadata, PC1Variance, PC2Variance, Groupcolors,miny, maxy, minx, maxx, pointSize, intgroup) {
  #Normalized (variance stabilized transformation), then run PCA
  vsd <- varianceStabilizingTransformation(dds, blind=TRUE)
  PCA<-plotPCA(vsd, intgroup=intgroup,returnData=TRUE)
  
  #plot
  p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=pointSize)+
  scale_color_manual(values=Groupcolors)+
  theme_classic()+
  ylim(miny,maxy)+
  xlim(minx,maxx)+
  ylab(paste("PC2: ", PC2Variance, "% Variance", sep =""))+
  xlab(paste("PC1: ", PC1Variance, "% Variance", sep =""))+
  theme(axis.text.y   = element_text(size=15, colour = "black", face="bold"),
        axis.title.y  = element_text(size=18, colour = "black", face="bold"),
        axis.title.x  = element_text(size=18, colour = "black", face="bold"),
        axis.text.x  = element_text(size=15, colour = "black", face="bold"),
        legend.background = element_rect(fill=NA, size=1, 
                                         linetype="solid", color="black"),
        legend.title=element_blank(),
        legend.text = element_text(size=18, face="bold"),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm"),
        legend.position = c(.85,.9),
        panel.border = element_rect(colour = "black", fill=NA, size=2)
        )
  return(p)
}

#Plot heatmap: takes as input DE results (padj ordered lowest to highest, rlog transformed data, metadata (blocker status and Sample type), top genes to include, and min and max subtract (used for heatmap aesthetics))
PlotHeatmap<- function(rld, metadata, topgenes, 
                       MinSubtract, MaxSubtract, fontsize, annotationColors){
  mat<- assay(rld)[topgenes,]
  mat<- mat - rowMeans(mat)
  colnames(mat)<-metadata$Sample
  
  #Heatmap settings
  mat_breaks<- seq(min(mat-MinSubtract), max(mat-MaxSubtract), length.out=75)

  #Create metadata- to be used to color code to show which group the sample belongs to
  metadata<- data.frame(Group=metadata$sampleGroup, 
                        row.names=metadata$Sample)

  #plot heatmap
  p<-pheatmap(mat, breaks = mat_breaks, 
            color =colorRampPalette( c("red", "black", "green"), space="rgb")(100),
            show_rownames = TRUE, show_colnames=TRUE,
            annotation = metadata,
            annotation_colors= annotationColors,
            fontsize_row = fontsize,fontsize=12,treeheight_row=0,
            border_color= NA,
            clustering_distance_cols = "euclidean")
  return(p)
}

#function that takes dds object and desired contrasts as input, outputs ordered results
ContrastDGE<- function(contrasts, dds){
  ContrastsUse<- c("sampleGroup", contrasts)
  ContrastsResults<- na.omit(results(dds, contrast=ContrastsUse))
  resOrdered<- ContrastsResults[order(ContrastsResults$padj),]
  return(resOrdered)
}

#Get sig DEGs based on 3 filters. Prints out # of DEGs fitting that criteria
GetSigDEGs<- function(DEGs, sigThresh, BasemeanThresh, Log2FCThresh){
        df<-DEGs[DEGs$padj<sigThresh & DEGs$baseMean>BasemeanThresh & abs(DEGs$log2FoldChange) > Log2FCThresh,]
        print(nrow(df))
        return(df)
}

#Run gsea using clusterprofiler
RunClusterProfilerGSEA<-function(DEGResults, gmtFile, pvalueThresh){
        #Create gene list: gene ID + adjusted p value
        DEGdf<-data.frame(geneID=rownames(DEGResults), 
                          padj=DEGResults$padj, 
                          log2FC=DEGResults$log2FoldChange)
        
        #Order genes by log2FC and convert to list format 
        DEGdf_ordered<-DEGdf[order(DEGdf$log2FC, decreasing=TRUE),]
        DEGlist_ordered<-DEGdf_ordered[,3]
        names(DEGlist_ordered)<-DEGdf_ordered$geneID
        
        #Read in desired .gmt file from clusterProfiler
        gmtfile <- system.file("extdata",gmtFile, package="clusterProfiler")
        gmt <- read.gmt(gmtfile)
        
        gseaResults <- GSEA(DEGlist_ordered, TERM2GENE=gmt, verbose=FALSE, pvalueCutoff = pvalueThresh)
        
        print(paste(nrow(gseaResults@result), "significant pathways identified"))
        return(gseaResults)
}

#Plot GSEA results in nice bar graph- from clusterprofiler analysis
PlotGSEABar<-function(GSEAResults, colors, yfontsize){
                ggplot(GSEAResults, aes(reorder(Description, NES), NES)) +
                geom_col(aes(fill=Group)) +
                coord_flip() +
                labs(x="Pathway", y="Normalized Enrichment Score") + 
                scale_fill_manual(values=colors)+
                theme_minimal()+
                theme(axis.title.x = element_text(size=14, color="black", face="bold"),
                      axis.title.y = element_blank(),
                      axis.text.x = element_text(size=12, color="black", face="bold"),
                      axis.text.y = element_text(size=yfontsize, color="black", face="bold"),
                      legend.title = element_blank(),
                      legend.text = element_text(size=yfontsize, color="black", face="bold"))
                
        }


#Run pathway analysis with gsea, plot- no longer in use
RunPlotGSEA<-function(res, gmt, nPermutations, OnlyPlotSig){
        #convert to vector
        ranks <- deframe(res)
        
        # Load the pathways into a named list
        pathways.hallmark <- gmtPathways(paste("C:/Users/Jenna/Documents/UKy_RNASeq/mSigdb/", gmt, sep=""))
        
        #run gsea with x permutations
        fgseaRes <- fgsea(pathways=pathways.hallmark, stats=ranks, nperm=nPermutations)
        
        #tidy up results
        fgseaResTidy <- fgseaRes %>%
        as_tibble() %>%
        arrange(desc(NES))
        
        #Only plot sig pathways, if desired
        if (OnlyPlotSig=="TRUE"){
                fgseaResTidy<-fgseaResTidy[fgseaResTidy$padj<0.05,]
        } else {
                fgseaResTidy<-fgseaResTidy
        }
                
        
        #plot results
        p<-ggplot(fgseaResTidy, aes(reorder(pathway, NES), NES)) +
        geom_col(aes(fill=padj<0.05)) +
        coord_flip() +
        labs(x="Pathway", y="Normalized Enrichment Score",
             title=paste("pathways from GSEA:", gmt, "/", nPermutations, " Permutations")) + 
        theme_minimal()
        
        return(list(p,as.data.frame(fgseaResTidy)))
}
```






Read in counts data (generated in"UKy_MergingRNACountsFilesAcross3Runs.R"), format, filter out very low counts nd "no_feature", and perform downsampling
```{r}
#Set wd
readwd<-"C:/Users/Jenna/OneDrive/Documents/UKy_RNASeq/RawCounts/"

#read in count data, reformat
countsInput<- read.csv(paste(readwd, "RawCounts_Run1S10.14_Run2Controls_AllOthersMerged.csv", sep =""), row.names="GeneID")

countsInput<-countsInput[,c(2:10, 19, 11:13,20,14:18, 21:25)]
```

```{r}
#remove sample 10 and 14- too low of counts- and sample 4- technical replicate of sample 2
counts<- countsInput[,!(colnames(countsInput)%in% c("X10", "X14", "X4"))]

#Remove highest expressed "gene" - no_feature. Could throw off DE
counts<-counts[!(rownames(counts) %in% "no_feature"),]

#Get rid of any genes with NA values, or where counts are 0 for that gene across all samples
counts<- na.omit(counts) #no NA values
counts<- counts[rowSums(counts)>0,] 

#downsample: set seed so results are replicable
counts<- downsample.counts(counts, seed=42) 
```

```{r}
detach("package:metaseqR")
detach("package:DESeq")
```


Read in metadata, reformat
```{r}
#Read in metadata info
wd<- "C:/Users/Jenna/Documents/UKy_RNASeq/"
meta<- read.csv(paste(wd, "MetaData_AllSamples.csv",sep=""))
rownames(meta)<- meta$Sample
meta<- meta[,-1]
meta<- meta[!(rownames(meta) %in% c(paste("X", c(10,14,4), sep=""))),] 
```


Run DESeq2, comparing Controls vs all CM. Plot PCA
```{r}
#Convert metadata to CM or Controls
meta_ControlsvCM<-meta
meta_ControlsvCM$sampleGroup<-gsub("TTNpos", "CM", meta_ControlsvCM$sampleGroup)
meta_ControlsvCM$sampleGroup<-gsub("TTNneg", "CM", meta_ControlsvCM$sampleGroup)

#RunDESeq2
dds_ControlsvCM<- RunDESeq2(counts, meta_ControlsvCM)

#Extract results
res_ControlsvCM<-as.data.frame(results(dds_ControlsvCM))

#Order results by pvalue
resOrdered_ControlsvCM<- res_ControlsvCM[order(res_ControlsvCM$padj),]

#Plot PCA- sampleGroup
MyPCA(dds_ControlsvCM, meta_ControlsvCM, 27,15, c("steelblue4","olivedrab"), -20,25,-30,20,5, "sampleGroup")

#Plot PCA- Sex
MyPCA(dds_ControlsvCM, meta_ControlsvCM, 27,15, c("violetred3","steelblue3"), -20,25,-30,20,5, "Sex")

#Plot PCA- Sex
MyPCA(dds_ControlsvCM, meta_ControlsvCM, 27,15, c("steelblue4","olivedrab"), -20,25,-30,20,5, "demo_race")
```

Plot heatmap for Controls vs CM analysis
```{r}
rld_ControlsVsCM<- rlog(dds_ControlsvCM)

#Select top sig DE mRNAs
topgenes<- rownames(resOrdered_ControlsvCM[1:80,])

#Coloring based on Controls/CM
annotationColors<-list(Group=c(Control="steelblue4", CM="olivedrab"))

PlotHeatmap(rld_ControlsVsCM,
            meta_ControlsvCM,
            topgenes,-.5,2.5,8,
            annotationColors)


#Coloring based on TTN+/TTN-
annotationColors<-list(Group=c(Control="steelblue4", TTNpos="goldenrod2", TTNneg="violetred4"))

PlotHeatmap(rld_ControlsVsCM,
            meta,
            topgenes,-.5,2.5,8,
            annotationColors)
```


#Run DESeq2 on downsampled counts, comparing TTN+/TTN-
```{r}
#DESeq2 on downsampled counts- pos vs neg vs controls
dds<- RunDESeq2(counts,meta)
```

Run and plot PCA for raw and DS counts
```{r}
MyPCA(dds, meta, 27,15, c("steelblue4","goldenrod2", "violetred4"), -20,25,-30,20,5, "sampleGroup")
```
```{r}
#Adjust metadata- move sample 3 back to TTN- group, plot by PCA
meta_sample3<-meta
meta_sample3[3,2]<- "TTNneg"

#DESeq2 on downsampled counts- pos vs neg vs controls
dds_sample3<- RunDESeq2(counts,meta_sample3)

MyPCA(dds_sample3, meta_sample3, 27,15, c("steelblue4","goldenrod2", "violetred4"), -20,25,-30,20,5, "sampleGroup")
```


```{r}
#Extract comparisons for 3 desired groups
DGERes_controlpos<- ContrastDGE(c("Control", "TTNpos"), dds)
DGERes_controlneg<- ContrastDGE(c("Control", "TTNneg"), dds)
DGERes_posneg<- ContrastDGE(c("TTNpos", "TTNneg"), dds)

nrow(DGERes_controlpos[DGERes_controlpos$padj<0.05,])
nrow(DGERes_controlneg[DGERes_controlneg$padj<0.05,])
nrow(DGERes_posneg[DGERes_posneg$padj<0.05,])
```

#Plot heatmap
```{r}
#Normalize
#rld<- rlog(dds)

#select top DE genes for control vs pos, control vs neg, and pos vs neg
topn<- 25
topgenes<- unique(c(rownames(DGERes_controlneg[1:topn,]), 
                    rownames(DGERes_controlpos[1:topn,]), 
                    rownames(DGERes_posneg[1:topn,])))


#Set annotation colors to use
annotationColors<-list(Group=c(Control="steelblue4", TTNpos="goldenrod2", TTNneg="violetred4"))

PlotHeatmap(rld,meta,topgenes,-.2,2.5,8, annotationColors)
```


#Performing random sample permutation to show (hopefully) that our groups give the highest number of DEGs
```{r}
AllN_DEGs<-list()
count=1
while (count<=20){
        #Randomly assign each sample to "pos" or "neg"
        CMSamples<-meta[meta$sampleGroup!="control",]
        CMSamples_Random<- CMSamples
        CMSamples_Random$sampleGroup<-sample(CMSamples$sampleGroup)
        
        #merge back with control samples
        meta_random<- rbind(CMSamples_Random, meta[meta$sampleGroup=="control",])
        
        #Perform DE using this new metadata
        dds_random<- RunDESeq2(counts,meta_random)
        
        #Extract comparisons for pos vs neg
        DEGs<-ContrastDGE(c("pos", "neg"), dds_random)
        SigDEGs<- DEGs[DEGs$padj<0.05,]
        AllN_DEGs[count]<-nrow(SigDEGs)
        
        print(count)
        count=count+1
}


```


#Get DEGs specific to + and -, and DEGs common to +/-
```{r}
#Get sig DEGs- 3 potential thresholds
sigThres<-0.05
BasemeanThresh<-0
Log2FCThresh<-0.0

tPosDEGs<- GetSigDEGs(DGERes_controlpos, sigThres, BasemeanThresh, Log2FCThresh) 
tNegDEGs<- GetSigDEGs(DGERes_controlneg, sigThres, BasemeanThresh, Log2FCThresh)

#Get DEGs specific to tPos or tNeg, or common to tPos/tNeg
tPosSpecific<-tPosDEGs[!(rownames(tPosDEGs) %in% rownames(tNegDEGs)),] 
nrow(tPosSpecific) #461

tNegSpecific<-tNegDEGs[!(rownames(tNegDEGs) %in% rownames(tPosDEGs)),] #2,361 (p<0.05) / 1,668 (p<0.01)
nrow(tNegSpecific) #2338

CommonDEGs<-tPosDEGs[rownames(tPosDEGs) %in% rownames(tNegDEGs),]
nrow(CommonDEGs) #2,303

venn<-list(TTNpos=rownames(tPosDEGs), TTNneg=rownames(tNegDEGs))
ggVennDiagram(venn, label="both",color="black",size=2)+scale_fill_gradient(low="gray90",high = "gray34")
```

For p<0.05 thresh: #samples in tPos/tNeg is split ~60/40. But the # of unique DEGs is split ~80/20. That is- we see a lot fewer tPos specific DEGs than we might expect, if the groups were totally comparable. Maybe suggests that the tPos are more similar to one another than the tNeg. Also can see this in the shear number of DEGs found in tPos/tNeg.

#Plotting heatmap of tPos and tNeg specific DEGs only
no longer in use
```{r}
#Select just top DEGs
Topn<-2000
topgenes<- unique(c(rownames(tPosSpecific[1:topn,]), rownames(tNegSpecific[1:topn,])))
#topgenes<-unique(rownames(tPosSpecific), rownames(tNegSpecific)) #Option to plot all

#normalize rld counts based on rowMeans
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

#settings for heatmap
subtractFromMin<- -.7
subtractFromMax<- 1.8
lengthOut<- 100
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)

#Create metadata- to be used to color code to show which group the sample belongs to
metadata<- as.data.frame(meta[,2])
rownames(metadata)<-rownames(meta)
colnames(metadata)<-"sampleGroup"

#plot heatmap
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)), show_rownames = FALSE, annotation_col = metadata, border_color=NA)
```


#For DEgs found in both tPos and tNeg- look at pattern of expression. Expect it to be very similar.
```{r}
#Separate counts for tPos and tNeg samples
tPosCounts<- as.data.frame(rowMeans(counts[,colnames(counts) %in% as.character(meta[meta$sampleGroup=="TTNpos",1])]))
tNegCounts<- as.data.frame(rowMeans(counts[,colnames(counts) %in% as.character(meta[meta$sampleGroup=="TTNneg",1])]))

#Merge counts into 1 df- for each DEG in both, mean raw counts for tPos and tNeg
CommonDEGs_counts<-merge(as.data.frame(CommonDEGs), tPosCounts, by=0)[,c(1,8)]
rownames(CommonDEGs_counts)<-CommonDEGs_counts$Row.names
CommonDEGs_counts<- merge(CommonDEGs_counts, tNegCounts, by=0)[,2:4]
colnames(CommonDEGs_counts)<-c("CommonDEGs", "Mean_tPos", "Mean_tNeg")
rownames(CommonDEGs_counts)<-CommonDEGs_counts$CommonDEGs
CommonDEGs_counts<- CommonDEGs_counts[,-1]

#Optional: transformation
CommonDEGs_log2<- log2(CommonDEGs_counts+0.1)
CommonDEGs_CPM<- (CommonDEGs_counts/colSums(counts)) * 1000000
CommonDEGs_CPM_log2<- log2(CommonDEGs_CPM+0.1)

df<-CommonDEGs_CPM_log2

p<-ggplot(df, aes(x=Mean_tPos, y=Mean_tNeg))+
        geom_point() +
        theme_classic()+
        xlab("Mean Log2(CPM) of Common DEGs: TTNpos samples")+
        ylab("Mean Log2(CPM) of Common DEGs: TTNneg samples")+
        geom_abline(slope=1, intercept = 0, color="red", size=1)+
        theme(axis.title.x = element_text(size=15, face="bold", color="black"),
              axis.title.y = element_text(size=15, face="bold", color="black"),
              axis.text.x = element_text(size=12, face="bold", color="black"),
              axis.text.y = element_text(size=12, face="bold", color="black"))
```



#Pathway analysis using fgsea
```{r}
#database options I've downloaded
PathwayOptions<-c("h.all.v7.1.symbols.gmt",
                  "c2.cp.kegg.v7.1.symbols.gmt",
                  "c3.all.v7.1.symbols.gmt",
                  "c5.all.v7.1.symbols.gmt")

#Reformat list of DEGs in tPos only: only need gene ID and stat
tPosPath<- as.data.frame(tPosSpecific[,c(2,4,6)])
tPosPath$geneID<-rownames(tPosPath)
tPosPathInput<-tPosPath[,c(4,2)]

#Reformat list of DEGs in tNeg only: only need gene ID and stat
tNegPath<- as.data.frame(tNegSpecific[,c(2,4,6)])
tNegPath$geneID<-rownames(tNegPath)
tNegPathInput<-tNegPath[,c(4,2)]

#Reformat list of DEGs in both tPos and tNeg: only need gene ID and stat
CommonDEGsPath<- as.data.frame(CommonDEGs[,c(2,4,6)])
CommonDEGsPath$geneID<-rownames(CommonDEGsPath)
CommonDEGsPathInput<-CommonDEGsPath[,c(4,2)]
        

#Running pathway analysis
gsea<-RunPlotGSEA(CommonDEGsPathInput, PathwayOptions[1], 10000, "TRUE")

#Hallmark database: saving for later
tNeg_Hallmark<-RunPlotGSEA(tNegPathInput, PathwayOptions[1], 10000, "TRUE")[[2]]
tPos_Hallmark<-RunPlotGSEA(tPosPathInput, PathwayOptions[1], 100000, "TRUE")[[2]]

#Running for all 4 databases. To combine and export tables.
CommonDEGs_SigPathways<-lapply(PathwayOptions, function(x){RunPlotGSEA(CommonDEGsPathInput, x, 10000, "TRUE")[[2]]})
all<-do.call("rbind", CommonDEGs_SigPathways) #110 pathways
all$leadingEdge<-as.character(all$leadingEdge)

tNeg_SigPathways<-lapply(PathwayOptions, function(x){RunPlotGSEA(tNegPathInput, x, 10000, "TRUE")[[2]]})
all<-do.call("rbind", tNeg_SigPathways) #517 pathways
all$leadingEdge<-as.character(all$leadingEdge)

write.csv(all, paste(wd, "DEGsCommonTo_BothtPostNeg_GSEAResults_FourPathwayDatabases_051320.csv",sep=""))

#Plot Hallmark pathways for tNeg and tPos on one plot
tPos_H<-tPos_Hallmark[,c(1,5,6)]
tNeg_H<-tNeg_Hallmark[,c(1,5,6)]

tPos_H$Group<-"tPos"
tNeg_H$Group<-"tNeg"
All_H<-rbind(tPos_H, tNeg_H)

#plot results
p<-ggplot(All_H, aes(reorder(pathway, NES), NES)) +
        geom_col(aes(fill=Group)) +
        coord_flip() +
        labs(x="Pathway", y="Normalized Enrichment Score") + 
        theme_minimal()
```



#Alternative: clusterProfiler
Several databases are available to perform GSEA analysis, including the Hallmark dataset, KEGG database, and GO terms.Here, the KEGG database will be used. The clusterProfiler package is also used to perform this analysis.



```{r, include=TRUE}
#Run clusterProfiler gsea- using Hallmark, KEGG, and GO databases for tPos specific, tNeg specific, and common DEGs

#Tpos specific
        gsea_tPosSpecific_Hallmark<-RunClusterProfilerGSEA(tPosSpecific, PathwayOptions[1], 0.5) #8
        gsea_tPosSpecific_KEGG<-RunClusterProfilerGSEA(tPosSpecific, PathwayOptions[2], 0.8) #3
        gsea_tPosSpecific_GO<-RunClusterProfilerGSEA(tPosSpecific, PathwayOptions[4], 0.05) #36

#Tneg specific
        gsea_tNegSpecific_Hallmark<-RunClusterProfilerGSEA(tNegSpecific, PathwayOptions[1], 0.1) #12
        gsea_tNegSpecific_KEGG<-RunClusterProfilerGSEA(tNegSpecific, PathwayOptions[2], 0.1) #6
        gsea_tNegSpecific_GO<-RunClusterProfilerGSEA(tNegSpecific, PathwayOptions[4], 0.001) #59
        
#Common DEGs
        gsea_CommonDEGs_Hallmark<-RunClusterProfilerGSEA(CommonDEGs, PathwayOptions[1], 0.1) #6
        gsea_CommonDEGs_KEGG<-RunClusterProfilerGSEA(CommonDEGs, PathwayOptions[2], 0.1) #4
        gsea_CommonDEGs_GO<-RunClusterProfilerGSEA(CommonDEGs, PathwayOptions[4], 0.001) #42

#tPos vs tNeg DE analysis
        PosvNegDEGs<- DGERes_posneg[DGERes_posneg$padj<0.1,] #Low threshold for more DEGs
        gsea_PosvNeg_Hallmark<-RunClusterProfilerGSEA(PosvNegDEGs, PathwayOptions[1], 0.5) #7
        gsea_PosvNeg_KEGG<-RunClusterProfilerGSEA(PosvNegDEGs, PathwayOptions[2], 0.5) #1
        gsea_PosvNeg_GO<-RunClusterProfilerGSEA(PosvNegDEGs, PathwayOptions[4], 0.3) #13

#Controls vs CM DE analysis
        ControlsvCMDEGs<-na.omit(res_ControlsvCM[res_ControlsvCM$padj<0.05,])
        gsea_ControlsvCM_Hallmark<-RunClusterProfilerGSEA(ControlsvCMDEGs, PathwayOptions[1], 0.05) #7
        gsea_ControlsvCM_KEGG<-RunClusterProfilerGSEA(ControlsvCMDEGs, PathwayOptions[2], 0.05) #7
        gsea_ControlsvCM_GO<-RunClusterProfilerGSEA(ControlsvCMDEGs, PathwayOptions[4], 0.0001) #80
        
        
#Plot Hallmark pathways for tNeg and tPos
        
        #Subset results
        tPos_H<-gsea_tPosSpecific_Hallmark@result[,c(2,5,7)]
        tNeg_H<-gsea_tNegSpecific_Hallmark@result[,c(2,5,7)]
        tPosvtNeg_H<-gsea_PosvNeg_Hallmark@result[,c(2,5,7)]
        Common_H<-gsea_CommonDEGs_Hallmark@result[,c(2,5,7)]
        ControlsvsCM_H<-gsea_ControlsvCM_Hallmark@result[,c(2,5,7)]
        
        #Add column for group, bind together into 1 df
        tPos_H$Group<-"TTNpos DEGs"
        tNeg_H$Group<-"TTNneg DEGs"
        tPosvtNeg_H$Group<- "tPos vs tNeg"
        Common_H$Group<-"Common DEGs"
        ControlsvsCM_H$Group<-"CM vs Controls"
        All_H<-rbind(tPos_H, tNeg_H, Common_H, ControlsvsCM_H)

        #plot results
        gseaColors<- c("steelblue4","olivedrab","goldenrod2", "violetred4")
        p_hallmark<-PlotGSEABar(All_H, gseaColors, 10)
       
#Plot KEGG pathways for tNeg and tPos
        
        #Subset results
        tPos_K<-gsea_tPosSpecific_KEGG@result[,c(2,5,7)]
        tNeg_K<-gsea_tNegSpecific_KEGG@result[,c(2,5,7)]
        tPosvtNeg_K<-gsea_PosvNeg_KEGG@result[,c(2,5,7)]
        Common_K<-gsea_CommonDEGs_KEGG@result[,c(2,5,7)]
        ControlsvsCM_K<-gsea_ControlsvCM_KEGG@result[,c(2,5,7)]
        
        #Add column for group, bind together into 1 df
        tPos_K$Group<-"TTNpos DEGs"
        tNeg_K$Group<-"TTNneg DEGs"
        tPosvtNeg_K$Group<- "tPos vs tNeg"
        Common_K$Group<-"Common DEGs"
        ControlsvsCM_K$Group<-"CM vs Controls"
        All_K<-rbind(tPos_K, tNeg_K, Common_K, ControlsvsCM_K)
        
        #plot results
        p_kegg<-PlotGSEABar(All_K, gseaColors, 10)

#Plot GO pathways for tNeg and tPos
        
        #Subset results
        tPos_G<-gsea_tPosSpecific_GO@result[abs(gsea_tPosSpecific_GO@result$NES)>2,c(2,5,7)]
        tNeg_G<-gsea_tNegSpecific_GO@result[abs(gsea_tNegSpecific_GO@result$NES)>2.4,c(2,5,7)]
        tPosvtNeg_G<-gsea_ControlsvCM_GO@result[abs(gsea_tPosSpecific_GO@result$NES)>2.1,c(2,5,7)]
        
        #Add column for group, bind together into 1 df
        tPos_G$Group<-"TTNpos DEGs"
        tNeg_G$Group<-"TTNneg DEGs"
        tPosvtNeg_G$Group<- "tPos vs tNeg"
        All_G<-rbind(tPos_G, tNeg_G)
        
        #plot results
        p_go<-PlotGSEABar(All_G, gseaColors, 7)

```



```{r}
#Plot Hallmark pathways for Controls vs CM
        
        #Subset results
        Common_H<-gsea_CommonDEGs_Hallmark@result[,c(2,5,7)]
        ControlsvsCM_H<-gsea_ControlsvCM_Hallmark@result[,c(2,5,7)]

        #Add column for group, bind together into 1 df
        Common_H$Group<-"Common DEGs"
        ControlsvsCM_H$Group<-"CM vs Controls"
        All_H<-rbind(Common_H, ControlsvsCM_H)

        #plot results
        gseaColors<- c("goldenrod2", "violetred4")
        p_hallmark<-PlotGSEABar(All_H, gseaColors, 10)
```





#Gene concept plotting
```{r}
p1 <- cnetplot(test, foldChange=DEGlist_ordered)
p3 <- cnetplot(test, foldChange=DEGlist_ordered, circular = TRUE, colorEdge = TRUE)
```

#Enrichment plot mapping
```{r}
p2 <- emapplot(test, pie_scale=1.5)
p3 <- emapplot(test, pie_scale=1.5,layout="kk")
p4 <- emapplot(test, pie_scale=1.5,layout="nicely") 
cowplot::plot_grid(p2, p3, p4, ncol=3, labels=LETTERS[1:3])
```


