---
title: "R Notebook"
output: html_notebook
---
#DE analysis of mRNA data

4 DE analyses performed:

1. Controls vs all CM
2. Controls vs TTN+
3. Controls vs TTN-
4. TTN+ vs TTN-


Import libraries:
```{r}
suppressPackageStartupMessages(library(DESeq2))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(pheatmap))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(metaseqR))
suppressPackageStartupMessages(library(ggVennDiagram))
suppressPackageStartupMessages(library(DT))
suppressPackageStartupMessages(library(tibble))
suppressPackageStartupMessages(library(fgsea))
suppressPackageStartupMessages(library(dplyr))
```


Functions
```{r}
#Create DESeq2 object, check that metadata is correct, and run DESeq2
RunDESeq2<- function(counts, meta){
  #check that counts/meta data has samples in same order
  print(all(rownames(meta) %in% colnames(counts)))
  print(all(rownames(meta) == colnames(counts)))
  
  #create DESeq2 object
  dds<- DESeqDataSetFromMatrix(countData=counts, 
                               colData=meta, 
                               design=~Sex + sampleGroup)
  
  #define levels- so that controls will always be used as reference group
  dds$sampleGroup<- relevel(dds$sampleGroup, ref= "Control")
  
  #run DE
  dds<- DESeq(dds)
  
  return(dds)

}


#Plot PCA nicely
MyPCA<- function(dds, metadata, PC1Variance, PC2Variance, Groupcolors,miny, maxy, minx, maxx, pointSize, intgroup) {
  #Normalized (variance stabilized transformation), then run PCA
  vsd <- varianceStabilizingTransformation(dds, blind=TRUE)
  PCA<-plotPCA(vsd, intgroup=intgroup,returnData=TRUE)
  
  #plot
  p<- ggplot(PCA, aes(x=PC1, y=PC2, color=group))+
  geom_point(size=pointSize)+
  scale_color_manual(values=Groupcolors)+
  theme_classic()+
  ylim(miny,maxy)+
  xlim(minx,maxx)+
  ylab(paste("PC2: ", PC2Variance, "% Variance", sep =""))+
  xlab(paste("PC1: ", PC1Variance, "% Variance", sep =""))+
  theme(axis.text.y   = element_text(size=15, colour = "black", face="bold"),
        axis.title.y  = element_text(size=18, colour = "black", face="bold"),
        axis.title.x  = element_text(size=18, colour = "black", face="bold"),
        axis.text.x  = element_text(size=15, colour = "black", face="bold"),
        legend.background = element_rect(fill=NA, size=1, 
                                         linetype="solid", color="black"),
        legend.title=element_blank(),
        legend.text = element_text(size=18, face="bold"),
        legend.key.size = unit(.4, "cm"),
        legend.spacing.y = unit(0, "mm"),
        legend.position = c(.85,.9),
        panel.border = element_rect(colour = "black", fill=NA, size=2)
        )
  return(p)
}

#Plot heatmap: takes as input DE results (padj ordered lowest to highest, rlog transformed data, metadata (blocker status and Sample type), top genes to include, and min and max subtract (used for heatmap aesthetics))
PlotHeatmap<- function(rld, metadata, topgenes, 
                       MinSubtract, MaxSubtract, fontsize, annotationColors){
  mat<- assay(rld)[topgenes,]
  mat<- mat - rowMeans(mat)
  colnames(mat)<-metadata$Sample
  
  #Heatmap settings
  mat_breaks<- seq(min(mat-MinSubtract), max(mat-MaxSubtract), length.out=75)

  #Create metadata- to be used to color code to show which group the sample belongs to
  metadata<- data.frame(Group=metadata$sampleGroup, 
                        row.names=metadata$Sample)

  #plot heatmap
  p<-pheatmap(mat, breaks = mat_breaks, 
            color =colorRampPalette( c("red", "black", "green"), space="rgb")(100),
            show_rownames = TRUE, show_colnames=TRUE,
            annotation = metadata,
            annotation_colors= annotationColors,
            fontsize_row = fontsize,fontsize=12,treeheight_row=0,
            border_color= NA,
            clustering_distance_cols = "euclidean")
  return(p)
}

#function that takes dds object and desired contrasts as input, outputs ordered results
ContrastDGE<- function(contrasts, dds){
  ContrastsUse<- c("sampleGroup", contrasts)
  ContrastsResults<- na.omit(results(dds, contrast=ContrastsUse))
  resOrdered<- ContrastsResults[order(ContrastsResults$padj),]
  return(resOrdered)
}

#Get sig DEGs based on 3 filters. Prints out # of DEGs fitting that criteria
GetSigDEGs<- function(DEGs, sigThresh, BasemeanThresh, Log2FCThresh){
        df<-DEGs[DEGs$padj<sigThresh & DEGs$baseMean>BasemeanThresh & abs(DEGs$log2FoldChange) > Log2FCThresh,]
        print(nrow(df))
        return(df)
}

```



Read in counts data (generated in"UKy_MergingRNACountsFilesAcross3Runs.R"), format, filter out very low counts nd "no_feature", and perform downsampling
```{r}
#Set wd
readwd<-"C:/Users/Jenna/OneDrive/Documents/UKy_RNASeq/RawCounts/"

#read in count data, reformat
countsInput<- read.csv(paste(readwd, "RawCounts_Run1S10.14_Run2Controls_AllOthersMerged.csv", sep =""), row.names="GeneID")

countsInput<-countsInput[,c(2:10, 19, 11:13,20,14:18, 21:25)]
```

Remove unused samples, no_feature, na, and downsample
```{r}
#remove sample 10 and 14- too low of counts- and sample 4- technical replicate of sample 2
counts<- countsInput[,!(colnames(countsInput)%in% c("X10", "X14", "X4"))]

#Remove highest expressed "gene" - no_feature. Could throw off DE
counts<-counts[!(rownames(counts) %in% "no_feature"),]

#Get rid of any genes with NA values, or where counts are 0 for that gene across all samples
counts<- na.omit(counts) #no NA values
counts<- counts[rowSums(counts)>0,] 

#downsample: set seed so results are replicable
counts<- downsample.counts(counts, seed=42) 
```

```{r}
detach("package:metaseqR")
detach("package:DESeq")
```


Read in metadata, reformat
```{r}
#Read in metadata info
wd<- "C:/Users/Jenna/Documents/UKy_RNASeq/"
meta<- read.csv(paste(wd, "MetaData_AllSamples.csv",sep=""))
rownames(meta)<- meta$Sample
meta<- meta[,-1]
meta<- meta[!(rownames(meta) %in% c(paste("X", c(10,14,4), sep=""))),] 
```


Run DESeq2, comparing Controls vs all CM. Plot PCA
```{r}
#Convert metadata to CM or Controls
meta_ControlsvCM<-meta
meta_ControlsvCM$sampleGroup<-gsub("TTNpos", "CM", meta_ControlsvCM$sampleGroup)
meta_ControlsvCM$sampleGroup<-gsub("TTNneg", "CM", meta_ControlsvCM$sampleGroup)

#RunDESeq2
dds_ControlsvCM<- RunDESeq2(counts, meta_ControlsvCM)

#Extract results
res_ControlsvCM<-as.data.frame(results(dds_ControlsvCM))

#Order results by pvalue
resOrdered_ControlsvCM<- res_ControlsvCM[order(res_ControlsvCM$padj),]

#Plot PCA- sampleGroup
MyPCA(dds_ControlsvCM, meta_ControlsvCM, 27,15, c("steelblue4","olivedrab"), -20,25,-30,20,5, "sampleGroup")

#Plot PCA- Sex
MyPCA(dds_ControlsvCM, meta_ControlsvCM, 27,15, c("violetred3","steelblue3"), -20,25,-30,20,5, "Sex")

#Plot PCA- Sex
MyPCA(dds_ControlsvCM, meta_ControlsvCM, 27,15, c("steelblue4","olivedrab"), -20,25,-30,20,5, "demo_race")
```

Plot heatmap for Controls vs CM analysis
```{r}
rld_ControlsVsCM<- rlog(dds_ControlsvCM)

#Select top sig DE mRNAs
topgenes<- rownames(resOrdered_ControlsvCM[1:80,])

#Coloring based on Controls/CM
annotationColors<-list(Group=c(Control="steelblue4", CM="olivedrab"))

PlotHeatmap(rld_ControlsVsCM,
            meta_ControlsvCM,
            topgenes,-.5,2.5,8,
            annotationColors)


#Coloring based on TTN+/TTN-
annotationColors<-list(Group=c(Control="steelblue4", TTNpos="goldenrod2", TTNneg="violetred4"))

PlotHeatmap(rld_ControlsVsCM,
            meta,
            topgenes,-.5,2.5,8,
            annotationColors)
```


#Run DESeq2 on downsampled counts, comparing TTN+/TTN-
```{r}
#DESeq2 on downsampled counts- pos vs neg vs controls
dds<- RunDESeq2(counts,meta)
```

Run and plot PCA for raw and DS counts
```{r}
MyPCA(dds, meta, 27,15, c("steelblue4","goldenrod2", "violetred4"), -20,25,-30,20,5, "sampleGroup")
```
```{r}
#Adjust metadata- move sample 3 back to TTN- group, plot by PCA
meta_sample3<-meta
meta_sample3[3,2]<- "TTNneg"

#DESeq2 on downsampled counts- pos vs neg vs controls
dds_sample3<- RunDESeq2(counts,meta_sample3)

MyPCA(dds_sample3, meta_sample3, 27,15, c("steelblue4","goldenrod2", "violetred4"), -20,25,-30,20,5, "sampleGroup")
```


```{r}
#Extract comparisons for 3 desired groups
DGERes_controlpos<- ContrastDGE(c("TTNpos","Control"), dds)
DGERes_controlneg<- ContrastDGE(c("TTNneg","Control"), dds)
DGERes_posneg<- ContrastDGE(c("TTNpos","TTNneg"), dds)

nrow(DGERes_controlpos[DGERes_controlpos$padj<0.05,])
nrow(DGERes_controlneg[DGERes_controlneg$padj<0.05,])
nrow(DGERes_posneg[DGERes_posneg$padj<0.05,])
```

#Plot heatmap
```{r}
#Normalize
rld<- rlog(dds)

#select top DE genes for control vs pos, control vs neg, and pos vs neg
topn<- 25
topgenes<- unique(c(rownames(DGERes_controlneg[1:topn,]), 
                    rownames(DGERes_controlpos[1:topn,]), 
                    rownames(DGERes_posneg[1:topn,])))


#Set annotation colors to use
annotationColors<-list(Group=c(Control="steelblue4", TTNpos="goldenrod2", TTNneg="violetred4"))

PlotHeatmap(rld,meta,topgenes,-.2,2.5,8, annotationColors)
```


#Performing random sample permutation to show (hopefully) that our groups give the highest number of DEGs
```{r}
AllN_DEGs<-list()
count=1
while (count<=20){
        #Randomly assign each sample to "pos" or "neg"
        CMSamples<-meta[meta$sampleGroup!="control",]
        CMSamples_Random<- CMSamples
        CMSamples_Random$sampleGroup<-sample(CMSamples$sampleGroup)
        
        #merge back with control samples
        meta_random<- rbind(CMSamples_Random, meta[meta$sampleGroup=="control",])
        
        #Perform DE using this new metadata
        dds_random<- RunDESeq2(counts,meta_random)
        
        #Extract comparisons for pos vs neg
        DEGs<-ContrastDGE(c("pos", "neg"), dds_random)
        SigDEGs<- DEGs[DEGs$padj<0.05,]
        AllN_DEGs[count]<-nrow(SigDEGs)
        
        print(count)
        count=count+1
}


```


#Get DEGs specific to + and -, and DEGs common to +/-
```{r}
#Get sig DEGs- 3 potential thresholds
sigThres<-0.05
BasemeanThresh<-0
Log2FCThresh<-0.0

tPosDEGs<- GetSigDEGs(DGERes_controlpos, sigThres, BasemeanThresh, Log2FCThresh) 
tNegDEGs<- GetSigDEGs(DGERes_controlneg, sigThres, BasemeanThresh, Log2FCThresh)

#Get DEGs specific to tPos or tNeg, or common to tPos/tNeg
tPosSpecific<-tPosDEGs[!(rownames(tPosDEGs) %in% rownames(tNegDEGs)),] 
nrow(tPosSpecific) #461

tNegSpecific<-tNegDEGs[!(rownames(tNegDEGs) %in% rownames(tPosDEGs)),] #2,361 (p<0.05) / 1,668 (p<0.01)
nrow(tNegSpecific) #2338

CommonDEGs<-tPosDEGs[rownames(tPosDEGs) %in% rownames(tNegDEGs),]
nrow(CommonDEGs) #2,303

venn<-list(TTNpos=rownames(tPosDEGs), TTNneg=rownames(tNegDEGs))
ggVennDiagram(venn, label="both",color="black",size=2)+scale_fill_gradient(low="gray90",high = "gray34")
```

For p<0.05 thresh: #samples in tPos/tNeg is split ~60/40. But the # of unique DEGs is split ~80/20. That is- we see a lot fewer tPos specific DEGs than we might expect, if the groups were totally comparable. Maybe suggests that the tPos are more similar to one another than the tNeg. Also can see this in the shear number of DEGs found in tPos/tNeg.

#Plotting heatmap of tPos and tNeg specific DEGs only
no longer in use
```{r}
#Select just top DEGs
Topn<-2000
topgenes<- unique(c(rownames(tPosSpecific[1:topn,]), rownames(tNegSpecific[1:topn,])))
#topgenes<-unique(rownames(tPosSpecific), rownames(tNegSpecific)) #Option to plot all

#normalize rld counts based on rowMeans
mat<- assay(rld)[topgenes,]
mat<- mat - rowMeans(mat)

#settings for heatmap
subtractFromMin<- -.7
subtractFromMax<- 1.8
lengthOut<- 100
mat_breaks<- seq(min(mat-subtractFromMin), max(mat-subtractFromMax), length.out=lengthOut)

#Create metadata- to be used to color code to show which group the sample belongs to
metadata<- as.data.frame(meta[,2])
rownames(metadata)<-rownames(meta)
colnames(metadata)<-"sampleGroup"

#plot heatmap
pheatmap(mat, breaks = mat_breaks, color =inferno(length(mat_breaks-1)), show_rownames = FALSE, annotation_col = metadata, border_color=NA)
```


#For DEgs found in both tPos and tNeg- look at pattern of expression. Expect it to be very similar.
```{r}
#Separate counts for tPos and tNeg samples
tPosCounts<- as.data.frame(rowMeans(counts[,colnames(counts) %in% as.character(meta[meta$sampleGroup=="TTNpos",1])]))
tNegCounts<- as.data.frame(rowMeans(counts[,colnames(counts) %in% as.character(meta[meta$sampleGroup=="TTNneg",1])]))

#Merge counts into 1 df- for each DEG in both, mean raw counts for tPos and tNeg
CommonDEGs_counts<-merge(as.data.frame(CommonDEGs), tPosCounts, by=0)[,c(1,8)]
rownames(CommonDEGs_counts)<-CommonDEGs_counts$Row.names
CommonDEGs_counts<- merge(CommonDEGs_counts, tNegCounts, by=0)[,2:4]
colnames(CommonDEGs_counts)<-c("CommonDEGs", "Mean_tPos", "Mean_tNeg")
rownames(CommonDEGs_counts)<-CommonDEGs_counts$CommonDEGs
CommonDEGs_counts<- CommonDEGs_counts[,-1]

#Optional: transformation
CommonDEGs_log2<- log2(CommonDEGs_counts+0.1)
CommonDEGs_CPM<- (CommonDEGs_counts/colSums(counts)) * 1000000
CommonDEGs_CPM_log2<- log2(CommonDEGs_CPM+0.1)

df<-CommonDEGs_CPM_log2

p<-ggplot(df, aes(x=Mean_tPos, y=Mean_tNeg))+
        geom_point() +
        theme_classic()+
        xlab("Mean Log2(CPM) of Common DEGs: TTNpos samples")+
        ylab("Mean Log2(CPM) of Common DEGs: TTNneg samples")+
        geom_abline(slope=1, intercept = 0, color="red", size=1)+
        theme(axis.title.x = element_text(size=15, face="bold", color="black"),
              axis.title.y = element_text(size=15, face="bold", color="black"),
              axis.text.x = element_text(size=12, face="bold", color="black"),
              axis.text.y = element_text(size=12, face="bold", color="black"))
```

Export all DE Results
```{r}
wd<-"C:/Users/Jenna/OneDrive/Documents/UKy/FilesInUse/"

write.csv(tPosSpecific, paste(wd, "DEResults_tPosSpecificSigDEGS.csv", sep=""))
write.csv(tNegSpecific, paste(wd, "DEResults_tNegSpecificSigDEGS.csv", sep=""))
write.csv(CommonDEGs, paste(wd, "DEResults_CommonSigDEGS.csv", sep=""))
write.csv(DGERes_posneg, paste(wd, "DEResults_PosvsNeg_AllGenes.csv",sep=""))
write.csv(resOrdered_ControlsvCM, paste(wd, "DEResults_ControlsvsCM_AllGenes.csv", sep=""))

```

```{r}
sessionInfo()
```


