---
title: "R Notebook"
output: html_notebook
---
#Perform GSEA pathway analysis on:

1. mRNA DE analysis results
2. miRNA DE results, following integartion with mRNA DE ("Integrating_RNASeq_and_miRNASeq_DEResults.Rmd")

##Two potential packages/methods considered:
1. clusterProfiler
2. fgsea

Both use GSEA algorithm. Main difference is that clusterProfiler allows for more extensive visualization. 

**Libraries**
```{r}
suppressPackageStartupMessages(library(clusterProfiler))
suppressPackageStartupMessages(library(fgsea))
suppressPackageStartupMessages(library(ggplot2))
```

**Functions**
```{r}
#Run gsea using clusterprofiler
RunClusterProfilerGSEA<-function(DEGResults, gmtFile, pvalueThresh){
        #Create gene list: gene ID + adjusted p value
        DEGdf<-data.frame(geneID=rownames(DEGResults), 
                          rank=DEGResults$rank)
        
        #Order genes by log2FC and convert to list format 
        DEGdf_ordered<-DEGdf[order(DEGdf$rank, decreasing=TRUE),]
        DEGlist_ordered<-DEGdf_ordered[,2]
        names(DEGlist_ordered)<-DEGdf_ordered$geneID
        
        #Read in desired .gmt file from clusterProfiler
        gmtfile <- system.file("extdata",gmtFile, package="clusterProfiler")
        gmt <- read.gmt(gmtfile)
        
        gseaResults <- GSEA(DEGlist_ordered, TERM2GENE=gmt, verbose=FALSE, pvalueCutoff = pvalueThresh)
        
        print(paste(nrow(gseaResults@result), "significant pathways identified"))
        return(gseaResults)
}

#Plot GSEA results in nice bar graph- from clusterprofiler analysis
 PlotGSEABar<-function(GSEAResults, colors, yfontsize, barsize, bardistance){
                ggplot(GSEAResults, aes(x=Description, y=NES)) +
                geom_bar(aes(fill=Group),width=barsize, 
                         position = position_dodge2(width = bardistance, preserve = "single"), 
                         stat = "identity") +
                coord_flip() +
                labs(x="Pathway", y="Normalized Enrichment Score") + 
                scale_fill_manual(values=colors)+
                theme_bw()+
                geom_hline(yintercept=0, color="grey4",size=1)+
                theme(axis.title.x = element_text(size=18, color="black", face="bold"),
                      axis.title.y = element_blank(),
                      axis.text.x = element_text(size=15, color="black", face="bold"),
                      axis.text.y = element_text(size=yfontsize, color="black", face="bold"),
                      legend.title = element_blank(),
                      legend.text = element_text(size=18, color="black", face="bold"),
                      panel.grid.major = element_line(colour="grey", size=.75),
                      panel.border = element_rect(colour = "black", fill=NA, size=3))
                
        } 


#Run pathway analysis with gsea, plot- no longer in use
RunPlotGSEA<-function(res, gmt, nPermutations, OnlyPlotSig){
        #convert to vector
        ranks <- deframe(res)
        
        # Load the pathways into a named list
        pathways.hallmark <- gmtPathways(paste("C:/Users/Jenna/Documents/UKy_RNASeq/mSigdb/", gmt, sep=""))
        
        #run gsea with x permutations
        fgseaRes <- fgsea(pathways=pathways.hallmark, stats=ranks, nperm=nPermutations)
        
        #tidy up results
        fgseaResTidy <- fgseaRes %>%
        as_tibble() %>%
        arrange(desc(NES))
        
        #Only plot sig pathways, if desired
        if (OnlyPlotSig=="TRUE"){
                fgseaResTidy<-fgseaResTidy[fgseaResTidy$padj<0.05,]
        } else {
                fgseaResTidy<-fgseaResTidy
        }
                
        
        #plot results
        p<-ggplot(fgseaResTidy, aes(reorder(pathway, NES), NES)) +
        geom_col(aes(fill=padj<0.05)) +
        coord_flip() +
        labs(x="Pathway", y="Normalized Enrichment Score",
             title=paste("pathways from GSEA:", gmt, "/", nPermutations, " Permutations")) + 
        theme_minimal()
        
        return(list(p,as.data.frame(fgseaResTidy)))
}
```

#Pathway Analysis for mRNA DE Results

**Read in DE results**
```{r}
wd<-"C:/Users/jenna/OneDrive/Documents/UKy/FilesInUse/"

tPosSpecific<-read.csv(paste(wd, "DEResults_tPosSpecificSigDEGS.csv", sep=""), row.names = "X")
tNegSpecific<-read.csv(paste(wd, "DEResults_tNegSpecificSigDEGS.csv", sep=""), row.names = "X")
CommonDEGs<- read.csv(paste(wd, "DEResults_CommonSigDEGS.csv", sep=""), row.names = "X")
DGERes_posneg<-read.csv(paste(wd, "DEResults_PosvsNeg_AllGenes.csv",sep=""), row.names = "X")
res_ControlsvCM<- read.csv(paste(wd, "DEResults_ControlsvsCM_AllGenes.csv", sep=""), row.names = "X")
```

**Calculate rank**
```{r}
tPosSpecific$rank<-as.numeric(gsub("FALSE", -1, gsub("TRUE", 1, tPosSpecific$log2FoldChange>0))) * -log10(tPosSpecific$padj)
tNegSpecific$rank<-as.numeric(gsub("FALSE", -1, gsub("TRUE", 1, tNegSpecific$log2FoldChange>0))) * -log10(tNegSpecific$padj)
CommonDEGs$rank<-as.numeric(gsub("FALSE", -1, gsub("TRUE", 1, CommonDEGs$log2FoldChange>0))) * -log10(CommonDEGs$padj)
DGERes_posneg$rank<-as.numeric(gsub("FALSE", -1, gsub("TRUE", 1, DGERes_posneg$log2FoldChange>0))) * -log10(DGERes_posneg$padj)
res_ControlsvCM$rank<-as.numeric(gsub("FALSE", -1, gsub("TRUE", 1, res_ControlsvCM$log2FoldChange>0))) * -log10(res_ControlsvCM$padj)
```


##clusterProfiler

Several databases are available to perform GSEA analysis, including the Hallmark dataset, KEGG database, and GO terms.

**Run GSEA analysis using all 3 databases for all 5 DEG lists**
```{r, include=TRUE}
PathwayOptions<-c("h.all.v7.1.symbols.gmt",
                  "c2.cp.kegg.v7.1.symbols.gmt",
                  "c5.all.v7.1.symbols.gmt")

#Run clusterProfiler gsea- using Hallmark, KEGG, and GO databases for tPos specific, tNeg specific, and common DEGs
set.seed(42)
#Tpos specific
        gsea_tPosSpecific_Hallmark<-RunClusterProfilerGSEA(tPosSpecific, PathwayOptions[1], 0.1) #8
        gsea_tPosSpecific_KEGG<-RunClusterProfilerGSEA(tPosSpecific, PathwayOptions[2], 0.1) #3
        gsea_tPosSpecific_GO<-RunClusterProfilerGSEA(tPosSpecific, PathwayOptions[3], 0.01) #13

#Tneg specific
        gsea_tNegSpecific_Hallmark<-RunClusterProfilerGSEA(tNegSpecific, PathwayOptions[1], 0.01) #12
        gsea_tNegSpecific_KEGG<-RunClusterProfilerGSEA(tNegSpecific, PathwayOptions[2], 0.01) #6
        gsea_tNegSpecific_GO<-RunClusterProfilerGSEA(tNegSpecific, PathwayOptions[3], 0.001) #57
        
#Common DEGs
        gsea_CommonDEGs_Hallmark<-RunClusterProfilerGSEA(CommonDEGs, PathwayOptions[1], 0.1) #6
        gsea_CommonDEGs_KEGG<-RunClusterProfilerGSEA(CommonDEGs, PathwayOptions[2], 0.1) #4
        gsea_CommonDEGs_GO<-RunClusterProfilerGSEA(CommonDEGs, PathwayOptions[3], 0.001) #42

#tPos vs tNeg DE analysis
        PosvNegDEGs<- DGERes_posneg[DGERes_posneg$padj<0.1,] #Low threshold for more DEGs
        gsea_PosvNeg_Hallmark<-RunClusterProfilerGSEA(PosvNegDEGs, PathwayOptions[1], 0.5) #7
        gsea_PosvNeg_KEGG<-RunClusterProfilerGSEA(PosvNegDEGs, PathwayOptions[2], 0.5) #1
        gsea_PosvNeg_GO<-RunClusterProfilerGSEA(PosvNegDEGs, PathwayOptions[3], 0.3) #13

#Controls vs CM DE analysis
        ControlsvCMDEGs<-na.omit(res_ControlsvCM[res_ControlsvCM$padj<0.05,])
        gsea_ControlsvCM_Hallmark<-RunClusterProfilerGSEA(ControlsvCMDEGs, PathwayOptions[1], 0.05) #7
        gsea_ControlsvCM_KEGG<-RunClusterProfilerGSEA(ControlsvCMDEGs, PathwayOptions[2], 0.05) #7
        gsea_ControlsvCM_GO<-RunClusterProfilerGSEA(ControlsvCMDEGs, PathwayOptions[3], 0.0001) #80
```


##Plot results

**Plot Hallmark pathways**
```{r}
#Subset results
tPos_H<-gsea_tPosSpecific_Hallmark@result[,c(2,5,7)]
tNeg_H<-gsea_tNegSpecific_Hallmark@result[,c(2,5,7)]
tPosvtNeg_H<-gsea_PosvNeg_Hallmark@result[,c(2,5,7)]
Common_H<-gsea_CommonDEGs_Hallmark@result[,c(2,5,7)]
ControlsvsCM_H<-gsea_ControlsvCM_Hallmark@result[,c(2,5,7)]

#Add column for group
tPos_H$Group<-"TTNpos DEGs"
tNeg_H$Group<-"TTNneg DEGs"
tPosvtNeg_H$Group<- "tPos vs tNeg"
Common_H$Group<-"Common DEGs"
ControlsvsCM_H$Group<-"CM vs Controls"

#split into 2 panels, then merge together so all descriptions are retained
panel1_H<-rbind(Common_H, ControlsvsCM_H)
panel2_H<-rbind(tPos_H, tNeg_H)
all_H<-merge(panel1_H, panel2_H, by="Description",all=TRUE)

#rename descriptors- remove Hallmark and "_"
all_H$Description<- gsub("HALLMARK", "", gsub("_", " ", all_H$Description))

#Convert NAs to 0
all_H[is.na(all_H)]<-0

#Convert 0s in Group to actual group- doesn't matter which one, won't be on graph
all_H$Group.x<-gsub(0,"Common DEGs", all_H$Group.x)
all_H$Group.y<-gsub(0, "TTNpos DEGs", all_H$Group.y)

#Split back into 2 panels
panel1_H<- all_H[,c(1:4)]
colnames(panel1_H)<-c("Description", "NES", "p.adjust", "Group")
panel2_H<-all_H[,c(1,5:7)]
colnames(panel2_H)<-c("Description", "NES", "p.adjust", "Group")


#plot results
gseaColors_panel1<- c("steelblue4","grey")
gseaColors_panel2<-c("goldenrod2", "violetred4")
p_hallmark_panel1<-PlotGSEABar(panel1_H, gseaColors_panel1, 14, .8,8)
p_hallmark_panel2<-PlotGSEABar(panel2_H, gseaColors_panel2, 14, .8,8)
grid.arrange(p_hallmark_panel1, p_hallmark_panel2, nrow=1)

```


**Plot KEGG pathways for tNeg and tPos**
```{r}
#Subset results
tPos_K<-gsea_tPosSpecific_KEGG@result[,c(2,5,7)]
tNeg_K<-gsea_tNegSpecific_KEGG@result[,c(2,5,7)]
tPosvtNeg_K<-gsea_PosvNeg_KEGG@result[,c(2,5,7)]
Common_K<-gsea_CommonDEGs_KEGG@result[,c(2,5,7)]
ControlsvsCM_K<-gsea_ControlsvCM_KEGG@result[,c(2,5,7)]

#Add column for group, bind together into 1 df
tPos_K$Group<-"TTNpos DEGs"
tNeg_K$Group<-"TTNneg DEGs"
tPosvtNeg_K$Group<- "tPos vs tNeg"
Common_K$Group<-"Common DEGs"
ControlsvsCM_K$Group<-"CM vs Controls"
All_K<-rbind(tPos_K, tNeg_K, Common_K, ControlsvsCM_K)

#split into 2 panels, then merge together so all descriptions are retained
panel1_K<-rbind(Common_K, ControlsvsCM_K)
panel2_K<-rbind(tPos_K, tNeg_K)
all_K<-merge(panel1_K, panel2_K, by="Description",all=TRUE)

#rename descriptors- remove KEGG and "_"
all_K$Description<- gsub("KEGG", "", gsub("_", " ", all_K$Description))

#Convert NAs to 0
all_K[is.na(all_K)]<-0

#Convert 0s in Group to actual group- doesn't matter which one, won't be on graph
all_K$Group.x<-gsub(0,"Common DEGs", all_K$Group.x)
all_K$Group.y<-gsub(0, "TTNpos DEGs", all_K$Group.y)

#Split back into 2 panels
panel1_K<- all_K[,c(1:4)]
colnames(panel1_K)<-c("Description", "NES", "p.adjust", "Group")
panel2_K<-all_K[,c(1,5:7)]
colnames(panel2_K)<-c("Description", "NES", "p.adjust", "Group")


#plot results
gseaColors_panel1<- c("steelblue4","grey")
gseaColors_panel2<-c("goldenrod2", "violetred4")
p_kegg_panel1<-PlotGSEABar(panel1_K, gseaColors_panel1, 14, .8,8)
p_kegg_panel2<-PlotGSEABar(panel2_K, gseaColors_panel2, 14, .8,8)
grid.arrange(p_kegg_panel1, p_kegg_panel2, nrow=1)

```


**Plot GO pathways for tNeg and tPos**
```{r}
#Subset results
tPos_G<-gsea_tPosSpecific_GO@result[order(abs(gsea_tPosSpecific_GO@result$NES), 
                                          decreasing = TRUE),c(2,5,7)]
tNeg_G<-gsea_tNegSpecific_GO@result[order(abs(gsea_tNegSpecific_GO@result$NES), 
                                          decreasing = TRUE),c(2,5,7)]
tPosvtNeg_G<-gsea_ControlsvCM_GO@result[order(abs(gsea_tPosSpecific_GO@result$NES),
                                              decreasing=TRUE),c(2,5,7)]
Common_G<-gsea_CommonDEGs_GO@result[order(abs(gsea_tPosSpecific_GO@result$NES), 
                                          decreasing=TRUE),c(2,5,7)]
ControlsvsCM_G<-gsea_ControlsvCM_GO@result[order(abs(gsea_tPosSpecific_GO@result$NES),
                                                 decreasing=TRUE),c(2,5,7)]

#Add column for group
tPos_G$Group<-"TTNpos DEGs"
tNeg_G$Group<-"TTNneg DEGs"
tPosvtNeg_G$Group<- "tPos vs tNeg"
Common_G$Group<-"Common DEGs"
ControlsvsCM_G$Group<-"CM vs Controls"

#split into 2 panels, then merge together so all descriptions are retained
panel1_G<-rbind(Common_G[1:8,], ControlsvsCM_G[1:8,])
panel2_G<-rbind(tPos_G[1:8,], tNeg_G[1:8,])
all_G<-merge(panel1_G, panel2_G, by="Description",all=TRUE)

#rename descriptors- remove GO and "_"
all_G$Description<- gsub("GO", "", gsub("_", " ", all_G$Description))

#Convert NAs to 0
all_G[is.na(all_G)]<-0

#Convert 0s in Group to actual group- doesn't matter which one, won't be on graph
all_G$Group.x<-gsub(0,"Common DEGs", all_G$Group.x)
all_G$Group.y<-gsub(0, "TTNpos DEGs", all_G$Group.y)

#Split back into 2 panels
panel1_G<- all_G[,c(1:4)]
colnames(panel1_G)<-c("Description", "NES", "p.adjust", "Group")
panel2_G<-all_G[,c(1,5:7)]
colnames(panel2_G)<-c("Description", "NES", "p.adjust", "Group")


#plot results
gseaColors_panel1<- c("steelblue4","grey")
gseaColors_panel2<-c("goldenrod2", "violetred4")
p_go_panel1<-PlotGSEABar(panel1_G, gseaColors_panel1, 14, .8,8)
p_go_panel2<-PlotGSEABar(panel2_G, gseaColors_panel2, 14, .8,8)
grid.arrange(p_go_panel1, p_go_panel2, nrow=1)

```


##Gene concept plotting
```{r}
DEGdf_ordered<-tNegSpecific[order(tNegSpecific$log2FoldChange, decreasing=TRUE),]
DEGlist_ordered<-DEGdf_ordered[,3]
names(DEGlist_ordered)<-rownames(DEGdf_ordered)

p1 <- cnetplot(gsea_tNegSpecific_Hallmark, foldChange=DEGlist_ordered)
p2 <- cnetplot(gsea_tNegSpecific_Hallmark, foldChange=DEGlist_ordered, circular = TRUE, colorEdge = TRUE,, node_label="none")
p2.5 <- cnetplot(gsea_tNegSpecific_Hallmark, node_label="none") 
```

##Enrichment plot mapping
```{r}
p3 <- emapplot(gsea_tNegSpecific_Hallmark, pie_scale=1.5)
p4 <- emapplot(gsea_ControlsvCM_GO, pie_scale=1.5,layout="kk")
p5 <- emapplot(gsea_ControlsvCM_GO, pie_scale=3,layout="nicely") 
```




#Pathway Analysis: miRNA/mRNA integration

**Read in results**
```{r}
wd<- "//Cifs2/rcdata$/UKRNASeq/"
miRTargets<-read.csv(paste(wd, "DESeq2_miRNA/multiMiR/multimirResults_log2padj.csv", sep =""), row.names="X")
```

**Calculate Rank and reformat**
```{r}
#Calculate rank value
FCSign<- gsub("FALSE", -1, gsub("TRUE", 1, miRTargets$mRNA_log2FC>0))
miRTargets$rank<- as.numeric(FCSign) * -log10(miRTargets$mRNA_padj)

#Selected needed columns + remove duplicated targets
GSEAInput_MiRTargets<-miRTargets[!duplicated(miRTargets$target_symbol),
                                         colnames(miRTargets) %in% c("target_symbol", "rank")]

#Rename columns
colnames(GSEAInput_MiRTargets)<-c("geneID", "rank")

#Rename rows
rownames(GSEAInput_MiRTargets)<- GSEAInput_MiRTargets$geneID

```


**Run GSEA**
```{r}
PathwayOptions<-c("h.all.v7.1.symbols.gmt",
                  "c2.cp.kegg.v7.1.symbols.gmt",
                  "c5.all.v7.1.symbols.gmt")

#Run clusterProfiler gsea- using Hallmark, KEGG, and GO databases for tPos specific, tNeg specific, and common DEGs
set.seed(42)


gsea_MiRTargets_Hallmark<-RunClusterProfilerGSEA(GSEAInput_MiRTargets, PathwayOptions[1], 0.05) #8
gsea_MiRTargets_KEGG<-RunClusterProfilerGSEA(GSEAInput_MiRTargets, PathwayOptions[2], 0.05) #3
gsea_MiRTargets_GO<-RunClusterProfilerGSEA(GSEAInput_MiRTargets, PathwayOptions[3], 0.01) #13

```


**Plot results**
```{r}
#Plot Hallmark pathways 
       
        #Subset results
        MiRTargets_H<-gsea_MiRTargets_Hallmark@result[,c(2,5,7)]
        ControlsvsCM_H<-gsea_ControlsvCM_Hallmark@result[,c(2,5,7)]
        
        #Add column for group, bind together with Controls vs CM pathways 
        MiRTargets_H$Group<-"DE mRNA targets of DE miRs"
        ControlsvsCM_H$Group<-"CM vs Controls"
        All_H<-rbind(MiRTargets_H, ControlsvsCM_H)

        #rename descriptors- remove Hallmark and "_"
        All_H$Description<- gsub("HALLMARK", "", gsub("_", " ", All_H$Description))
        
        #plot results
        gseaColors<- c("steelblue4", "olivedrab")
        p_hallmark<-PlotGSEABar(All_H, gseaColors, 14, 1,8)
       
#Plot KEGG pathways 
        
        #Subset results
        MiRTargets_K<-gsea_MiRTargets_KEGG@result[,c(2,5,7)]
        ControlsvsCM_K<-gsea_ControlsvCM_KEGG@result[,c(2,5,7)]
        
        #Add column for group, bind together with Controls vs CM pathways 
        MiRTargets_K$Group<-"DE mRNA targets of DE miRs"
        ControlsvsCM_K$Group<-"CM vs Controls"
        All_K<-rbind(MiRTargets_K, ControlsvsCM_K)
        
        #rename descriptors- remove Hallmark and "_"
        All_K$Description<- gsub("KEGG", "", gsub("_", " ", All_K$Description))

        #plot results
        gseaColors<- c("steelblue4", "olivedrab")
        p_KEGG<-PlotGSEABar(All_K, gseaColors, 17, 1,8)
 
               
#Plot GO pathways 
        
        #Subset results
        MiRTargets_G<-gsea_MiRTargets_GO@result[,c(2,5,7)]
        ControlsvsCM_G<-gsea_ControlsvCM_GO@result[,c(2,5,7)]
        
        #Add column for group, bind together with Controls vs CM pathways 
        MiRTargets_G$Group<-"DE mRNA targets of DE miRs"
        ControlsvsCM_G$Group<-"CM vs Controls"
        All_G<-rbind(MiRTargets_G, ControlsvsCM_G)
        
        #rename descriptors- remove Hallmark and "_"
        All_G$Description<- gsub("GO", "", gsub("_", " ", All_G$Description))

        #plot results
        gseaColors<- c("steelblue4", "olivedrab")
        p_GO<-PlotGSEABar(All_G, gseaColors, 14, 1,8)
       
```


##Pathway analysis using fgsea
No longer in use
```{r}
#database options I've downloaded
PathwayOptions<-c("h.all.v7.1.symbols.gmt",
                  "c2.cp.kegg.v7.1.symbols.gmt",
                  "c5.all.v7.1.symbols.gmt")

#Reformat list of DEGs in tPos only: only need gene ID and stat
tPosPath<- as.data.frame(tPosSpecific[,c(2,4,6)])
tPosPath$geneID<-rownames(tPosPath)
tPosPathInput<-tPosPath[,c(4,2)]

#Reformat list of DEGs in tNeg only: only need gene ID and stat
tNegPath<- as.data.frame(tNegSpecific[,c(2,4,6)])
tNegPath$geneID<-rownames(tNegPath)
tNegPathInput<-tNegPath[,c(4,2)]

#Reformat list of DEGs in both tPos and tNeg: only need gene ID and stat
CommonDEGsPath<- as.data.frame(CommonDEGs[,c(2,4,6)])
CommonDEGsPath$geneID<-rownames(CommonDEGsPath)
CommonDEGsPathInput<-CommonDEGsPath[,c(4,2)]
        

#Running pathway analysis
gsea<-RunPlotGSEA(CommonDEGsPathInput, PathwayOptions[1], 10000, "TRUE")

#Hallmark database: saving for later
tNeg_Hallmark<-RunPlotGSEA(tNegPathInput, PathwayOptions[1], 10000, "TRUE")[[2]]
tPos_Hallmark<-RunPlotGSEA(tPosPathInput, PathwayOptions[1], 100000, "TRUE")[[2]]

#Running for all 4 databases. To combine and export tables.
CommonDEGs_SigPathways<-lapply(PathwayOptions, function(x){RunPlotGSEA(CommonDEGsPathInput, x, 10000, "TRUE")[[2]]})
all<-do.call("rbind", CommonDEGs_SigPathways) #110 pathways
all$leadingEdge<-as.character(all$leadingEdge)

tNeg_SigPathways<-lapply(PathwayOptions, function(x){RunPlotGSEA(tNegPathInput, x, 10000, "TRUE")[[2]]})
all<-do.call("rbind", tNeg_SigPathways) #517 pathways
all$leadingEdge<-as.character(all$leadingEdge)

write.csv(all, paste(wd, "DEGsCommonTo_BothtPostNeg_GSEAResults_FourPathwayDatabases_051320.csv",sep=""))

#Plot Hallmark pathways for tNeg and tPos on one plot
tPos_H<-tPos_Hallmark[,c(1,5,6)]
tNeg_H<-tNeg_Hallmark[,c(1,5,6)]
tPos_H$Group<-"tPos"
tNeg_H$Group<-"tNeg"
All_H<-rbind(tPos_H, tNeg_H)

#plot results
p<-ggplot(All_H, aes(reorder(pathway, NES), NES)) +
        geom_col(aes(fill=Group)) +
        coord_flip() +
        labs(x="Pathway", y="Normalized Enrichment Score") + 
        theme_minimal()
```



```{r}
sessionInfo()
```

