---
title: "R Notebook"
output: html_notebook
---


Libraries
```{r}
suppressPackageStartupMessages(library(clusterProfiler))
suppressPackageStartupMessages(library(fgsea))
```

Functions
```{r}
#Run gsea using clusterprofiler
RunClusterProfilerGSEA<-function(DEGResults, gmtFile, pvalueThresh){
        #Create gene list: gene ID + adjusted p value
        DEGdf<-data.frame(geneID=rownames(DEGResults), 
                          padj=DEGResults$padj, 
                          log2FC=DEGResults$log2FoldChange)
        
        #Order genes by log2FC and convert to list format 
        DEGdf_ordered<-DEGdf[order(DEGdf$log2FC, decreasing=TRUE),]
        DEGlist_ordered<-DEGdf_ordered[,3]
        names(DEGlist_ordered)<-DEGdf_ordered$geneID
        
        #Read in desired .gmt file from clusterProfiler
        gmtfile <- system.file("extdata",gmtFile, package="clusterProfiler")
        gmt <- read.gmt(gmtfile)
        
        gseaResults <- GSEA(DEGlist_ordered, TERM2GENE=gmt, verbose=FALSE, pvalueCutoff = pvalueThresh)
        
        print(paste(nrow(gseaResults@result), "significant pathways identified"))
        return(gseaResults)
}

#Plot GSEA results in nice bar graph- from clusterprofiler analysis
PlotGSEABar<-function(GSEAResults, colors, yfontsize, barsize, bardistance){
                ggplot(GSEAResults, aes(reorder(Description, NES), NES)) +
                geom_bar(aes(fill=Group),width=barsize, 
                         position = position_dodge2(width = bardistance, preserve = "single"), 
                         stat = "identity") +
                coord_flip() +
                labs(x="Pathway", y="Normalized Enrichment Score") + 
                scale_fill_manual(values=colors)+
                theme_minimal()+
                theme(axis.title.x = element_text(size=14, color="black", face="bold"),
                      axis.title.y = element_blank(),
                      axis.text.x = element_text(size=12, color="black", face="bold"),
                      axis.text.y = element_text(size=yfontsize, color="black", face="bold"),
                      legend.title = element_blank(),
                      legend.text = element_text(size=yfontsize, color="black", face="bold"))
                
        }


#Run pathway analysis with gsea, plot- no longer in use
RunPlotGSEA<-function(res, gmt, nPermutations, OnlyPlotSig){
        #convert to vector
        ranks <- deframe(res)
        
        # Load the pathways into a named list
        pathways.hallmark <- gmtPathways(paste("C:/Users/Jenna/Documents/UKy_RNASeq/mSigdb/", gmt, sep=""))
        
        #run gsea with x permutations
        fgseaRes <- fgsea(pathways=pathways.hallmark, stats=ranks, nperm=nPermutations)
        
        #tidy up results
        fgseaResTidy <- fgseaRes %>%
        as_tibble() %>%
        arrange(desc(NES))
        
        #Only plot sig pathways, if desired
        if (OnlyPlotSig=="TRUE"){
                fgseaResTidy<-fgseaResTidy[fgseaResTidy$padj<0.05,]
        } else {
                fgseaResTidy<-fgseaResTidy
        }
                
        
        #plot results
        p<-ggplot(fgseaResTidy, aes(reorder(pathway, NES), NES)) +
        geom_col(aes(fill=padj<0.05)) +
        coord_flip() +
        labs(x="Pathway", y="Normalized Enrichment Score",
             title=paste("pathways from GSEA:", gmt, "/", nPermutations, " Permutations")) + 
        theme_minimal()
        
        return(list(p,as.data.frame(fgseaResTidy)))
}
```

Read in DE results
```{r}
wd<-"C:/Users/Jenna/OneDrive/Documents/UKy/FilesInUse/"

tPosSpecific<-read.csv(paste(wd, "DEResults_tPosSpecificSigDEGS.csv", sep=""), row.names = "X")
tNegSpecific<-read.csv(paste(wd, "DEResults_tNegSpecificSigDEGS.csv", sep=""), row.names = "X")
CommonDEGs<- read.csv(paste(wd, "DEResults_CommonSigDEGS.csv", sep=""), row.names = "X")
DGERes_posneg<-read.csv(paste(wd, "DEResults_PosvsNeg_AllGenes.csv",sep=""), row.names = "X")
res_ControlsvCM<- read.csv(paste(wd, "DEResults_ControlsvsCM_AllGenes.csv", sep=""), row.names = "X")
```



#clusterProfiler
Several databases are available to perform GSEA analysis, including the Hallmark dataset, KEGG database, and GO terms.Here, the KEGG database will be used. The clusterProfiler package is also used to perform this analysis.



```{r, include=TRUE}
PathwayOptions<-c("h.all.v7.1.symbols.gmt",
                  "c2.cp.kegg.v7.1.symbols.gmt",
                  "c5.all.v7.1.symbols.gmt")

#Run clusterProfiler gsea- using Hallmark, KEGG, and GO databases for tPos specific, tNeg specific, and common DEGs
set.seed(42)
#Tpos specific
        gsea_tPosSpecific_Hallmark<-RunClusterProfilerGSEA(tPosSpecific, PathwayOptions[1], 0.3) #8
        gsea_tPosSpecific_KEGG<-RunClusterProfilerGSEA(tPosSpecific, PathwayOptions[2], 0.3) #3
        gsea_tPosSpecific_GO<-RunClusterProfilerGSEA(tPosSpecific, PathwayOptions[3], 0.01) #13

#Tneg specific
        gsea_tNegSpecific_Hallmark<-RunClusterProfilerGSEA(tNegSpecific, PathwayOptions[1], 0.1) #12
        gsea_tNegSpecific_KEGG<-RunClusterProfilerGSEA(tNegSpecific, PathwayOptions[2], 0.1) #6
        gsea_tNegSpecific_GO<-RunClusterProfilerGSEA(tNegSpecific, PathwayOptions[3], 0.001) #57
        
#Common DEGs
        gsea_CommonDEGs_Hallmark<-RunClusterProfilerGSEA(CommonDEGs, PathwayOptions[1], 0.1) #6
        gsea_CommonDEGs_KEGG<-RunClusterProfilerGSEA(CommonDEGs, PathwayOptions[2], 0.1) #4
        gsea_CommonDEGs_GO<-RunClusterProfilerGSEA(CommonDEGs, PathwayOptions[3], 0.001) #42

#tPos vs tNeg DE analysis
        PosvNegDEGs<- DGERes_posneg[DGERes_posneg$padj<0.1,] #Low threshold for more DEGs
        gsea_PosvNeg_Hallmark<-RunClusterProfilerGSEA(PosvNegDEGs, PathwayOptions[1], 0.5) #7
        gsea_PosvNeg_KEGG<-RunClusterProfilerGSEA(PosvNegDEGs, PathwayOptions[2], 0.5) #1
        gsea_PosvNeg_GO<-RunClusterProfilerGSEA(PosvNegDEGs, PathwayOptions[3], 0.3) #13

#Controls vs CM DE analysis
        ControlsvCMDEGs<-na.omit(res_ControlsvCM[res_ControlsvCM$padj<0.05,])
        gsea_ControlsvCM_Hallmark<-RunClusterProfilerGSEA(ControlsvCMDEGs, PathwayOptions[1], 0.05) #7
        gsea_ControlsvCM_KEGG<-RunClusterProfilerGSEA(ControlsvCMDEGs, PathwayOptions[2], 0.05) #7
        gsea_ControlsvCM_GO<-RunClusterProfilerGSEA(ControlsvCMDEGs, PathwayOptions[3], 0.0001) #80
```

```{r}
#Plot Hallmark pathways for tNeg and tPos
        
        #Subset results
        tPos_H<-gsea_tPosSpecific_Hallmark@result[,c(2,5,7)]
        tNeg_H<-gsea_tNegSpecific_Hallmark@result[,c(2,5,7)]
        tPosvtNeg_H<-gsea_PosvNeg_Hallmark@result[,c(2,5,7)]
        Common_H<-gsea_CommonDEGs_Hallmark@result[,c(2,5,7)]
        ControlsvsCM_H<-gsea_ControlsvCM_Hallmark@result[,c(2,5,7)]
        
        #Add column for group, bind together into 1 df
        tPos_H$Group<-"TTNpos DEGs"
        tNeg_H$Group<-"TTNneg DEGs"
        tPosvtNeg_H$Group<- "tPos vs tNeg"
        Common_H$Group<-"Common DEGs"
        ControlsvsCM_H$Group<-"CM vs Controls"
        All_H<-rbind(tPos_H, tNeg_H, Common_H, ControlsvsCM_H, tPosvtNeg_H)

        #plot results
        gseaColors<- c("steelblue4","grey", "olivedrab", "violetred4", "goldenrod2")
        p_hallmark<-PlotGSEABar(All_H, gseaColors, 8, 1,8)
       
#Plot KEGG pathways for tNeg and tPos
        
        #Subset results
        tPos_K<-gsea_tPosSpecific_KEGG@result[,c(2,5,7)]
        tNeg_K<-gsea_tNegSpecific_KEGG@result[,c(2,5,7)]
        tPosvtNeg_K<-gsea_PosvNeg_KEGG@result[,c(2,5,7)]
        Common_K<-gsea_CommonDEGs_KEGG@result[,c(2,5,7)]
        ControlsvsCM_K<-gsea_ControlsvCM_KEGG@result[,c(2,5,7)]
        
        #Add column for group, bind together into 1 df
        tPos_K$Group<-"TTNpos DEGs"
        tNeg_K$Group<-"TTNneg DEGs"
        tPosvtNeg_K$Group<- "tPos vs tNeg"
        Common_K$Group<-"Common DEGs"
        ControlsvsCM_K$Group<-"CM vs Controls"
        All_K<-rbind(tPos_K, tNeg_K, Common_K, ControlsvsCM_K, tPosvtNeg_K)
        
        #plot results
        p_kegg<-PlotGSEABar(All_K, gseaColors, 12,1,2)

#Plot GO pathways for tNeg and tPos
        
        #Subset results
        tPos_G<-gsea_tPosSpecific_GO@result[order(abs(gsea_tPosSpecific_GO@result$NES), 
                                                  decreasing = TRUE),c(2,5,7)]
        tNeg_G<-gsea_tNegSpecific_GO@result[order(abs(gsea_tNegSpecific_GO@result$NES), 
                                                  decreasing = TRUE),c(2,5,7)]
        tPosvtNeg_G<-gsea_ControlsvCM_GO@result[order(abs(gsea_tPosSpecific_GO@result$NES),
                                                      decreasing=TRUE),c(2,5,7)]
        Common_G<-gsea_CommonDEGs_GO@result[order(abs(gsea_tPosSpecific_GO@result$NES), 
                                                  decreasing=TRUE),c(2,5,7)]
        ControlsvsCM_G<-gsea_ControlsvCM_GO@result[order(abs(gsea_tPosSpecific_GO@result$NES),
                                                         decreasing=TRUE),c(2,5,7)]
        
        #Add column for group, bind together into 1 df
        tPos_G$Group<-"TTNpos DEGs"
        tNeg_G$Group<-"TTNneg DEGs"
        tPosvtNeg_G$Group<- "tPos vs tNeg"
        Common_G$Group<-"Common DEGs"
        ControlsvsCM_G$Group<-"CM vs Controls"
        All_G<-rbind(tPos_G[1:8,], tNeg_G[1:8,], Common_G[1:8,], ControlsvsCM_G[1:8,],tPosvtNeg_G[1:8,])
        
        #plot results
        p_go<-PlotGSEABar(All_G, gseaColors, 7,1,2)

```



```{r}
common_ccp<-gsea_CommonDEGs_GO@result[gsea_CommonDEGs_GO@result$Description=="GO_CELL_CYCLE_PROCESS",]
ControlsvCM_ccp<-gsea_ControlsvCM_GO@result[gsea_ControlsvCM_GO@result$Description=="GO_CELL_CYCLE_PROCESS",]

common_genes<-unlist(as.list(strsplit(common_ccp$core_enrichment, '/')[[1]])) #all
controlsvCM_genes<-unlist(as.list(strsplit(ControlsvCM_ccp$core_enrichment, '/')[[1]])) #41 out of 88

common_counts<-CommonDEGs[rownames(CommonDEGs) %in% common_genes,]
controlsvCM_counts<-ControlsvCMDEGs[rownames(ControlsvCMDEGs) %in% common_genes,]
```





#Gene concept plotting
```{r}
DEGdf_ordered<-tNegSpecific[order(tNegSpecific$log2FoldChange, decreasing=TRUE),]
DEGlist_ordered<-DEGdf_ordered[,3]
names(DEGlist_ordered)<-rownames(DEGdf_ordered)

p1 <- cnetplot(gsea_tNegSpecific_Hallmark, foldChange=DEGlist_ordered)
p2 <- cnetplot(gsea_tNegSpecific_Hallmark, foldChange=DEGlist_ordered, circular = TRUE, colorEdge = TRUE)
```

#Enrichment plot mapping
```{r}
p3 <- emapplot(gsea_tNegSpecific_Hallmark, pie_scale=1.5)
p4 <- emapplot(gsea_ControlsvCM_GO, pie_scale=1.5,layout="kk")
p5 <- emapplot(gsea_ControlsvCM_GO, pie_scale=3,layout="nicely") 
```

No longer in use
#Pathway analysis using fgsea
```{r}
#database options I've downloaded
PathwayOptions<-c("h.all.v7.1.symbols.gmt",
                  "c2.cp.kegg.v7.1.symbols.gmt",
                  "c5.all.v7.1.symbols.gmt")

#Reformat list of DEGs in tPos only: only need gene ID and stat
tPosPath<- as.data.frame(tPosSpecific[,c(2,4,6)])
tPosPath$geneID<-rownames(tPosPath)
tPosPathInput<-tPosPath[,c(4,2)]

#Reformat list of DEGs in tNeg only: only need gene ID and stat
tNegPath<- as.data.frame(tNegSpecific[,c(2,4,6)])
tNegPath$geneID<-rownames(tNegPath)
tNegPathInput<-tNegPath[,c(4,2)]

#Reformat list of DEGs in both tPos and tNeg: only need gene ID and stat
CommonDEGsPath<- as.data.frame(CommonDEGs[,c(2,4,6)])
CommonDEGsPath$geneID<-rownames(CommonDEGsPath)
CommonDEGsPathInput<-CommonDEGsPath[,c(4,2)]
        

#Running pathway analysis
gsea<-RunPlotGSEA(CommonDEGsPathInput, PathwayOptions[1], 10000, "TRUE")

#Hallmark database: saving for later
tNeg_Hallmark<-RunPlotGSEA(tNegPathInput, PathwayOptions[1], 10000, "TRUE")[[2]]
tPos_Hallmark<-RunPlotGSEA(tPosPathInput, PathwayOptions[1], 100000, "TRUE")[[2]]

#Running for all 4 databases. To combine and export tables.
CommonDEGs_SigPathways<-lapply(PathwayOptions, function(x){RunPlotGSEA(CommonDEGsPathInput, x, 10000, "TRUE")[[2]]})
all<-do.call("rbind", CommonDEGs_SigPathways) #110 pathways
all$leadingEdge<-as.character(all$leadingEdge)

tNeg_SigPathways<-lapply(PathwayOptions, function(x){RunPlotGSEA(tNegPathInput, x, 10000, "TRUE")[[2]]})
all<-do.call("rbind", tNeg_SigPathways) #517 pathways
all$leadingEdge<-as.character(all$leadingEdge)

write.csv(all, paste(wd, "DEGsCommonTo_BothtPostNeg_GSEAResults_FourPathwayDatabases_051320.csv",sep=""))

#Plot Hallmark pathways for tNeg and tPos on one plot
tPos_H<-tPos_Hallmark[,c(1,5,6)]
tNeg_H<-tNeg_Hallmark[,c(1,5,6)]
tPos_H$Group<-"tPos"
tNeg_H$Group<-"tNeg"
All_H<-rbind(tPos_H, tNeg_H)

#plot results
p<-ggplot(All_H, aes(reorder(pathway, NES), NES)) +
        geom_col(aes(fill=Group)) +
        coord_flip() +
        labs(x="Pathway", y="Normalized Enrichment Score") + 
        theme_minimal()
```